### 2017-02-24 paper/main.tex
added serialiazation and paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
Compressed Directional Distance Transform for Fast 2D Ray Casting
}

\author{Corey H. Walsh$^{1}$ and Sertac Karaman$^{2}$% <-this % stops a space
% \thanks{*This work was not supported by any organization}% <-this % stops a space
\thanks{$^{1}$Corey H. Walsh is with the Department of Computer Science and Engineering, 
Albert Author is with Faculty of Electrical Engineering, Mathematics and Computer Science, Massachusetts Institute of Technology,
        Cambridge, MA 02139, USA
        {\tt\small chwalsh@mit.edu}}%
\thanks{$^{2}$Sertac Karaman is with Faculty of Computer Science and Engineering, Massachusetts Institute of Technology,
        Cambridge, MA 02139, USA
        {\tt\small sertac@mit.edu}}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

The ray cast operation is an important sub-problem for robot localization using distance sensors to detect the environment. In particle filter localization, particles are re-weighted according to a Bayesian update equation which relies on ray casting to determine the ground truth distance between the hypothesis pose and obstacles in a map. Effective particle filters often maintain thousands of particles, each of which must be reweighted tens of times per second, and each re-weighting may require tens of ray casts. As a result, millions of ray cast operations may be resolved per second, posing a significant computational challenge for resources constrained systems.

To combat this challenge, Fox et al. [1999] suggest the use of a large three-dimensional lookup table (LUT) to store expected ranges for each discrete (x,y,Θ) state. While this is simple to implement and does result in large speed improvements as compared to ray casting, it can be prohibitively memory intensive for large maps and/or resource constrained systems. We propose an analogous data structure called a Compressed Directional Distance Transform (CDDT) which requires significantly less memory and precomputation time while still allowing near constant time ray casting queries.

Many robotic applications require three dimensional maps, however that is considered outside the scope of this work since many data structures exist for accelerating 3D ray casting, such as the bounded volume hierarchy and related variants. The five dimensional table required to precompute expected ranges for a discrete state space (x,y,z,Θ,Φ) would be prohibitively expensive to both compute and store, so 3D acceleration data structures are exclusively used when 3D ray casting is required. Though our approach does extend to higher dimensionality and would be smaller than the analogous five dimensional table, it would still be prohibitively expensive in greater than two dimensions, thus we do not consider it here.

\section{RELATED WORK}

Several widely used algorithms exist for ray casting in a two dimensional space. Due to varied handling of edge cases and ambiguities in the problem of determining distance between any point and the nearest obstacle in a particular direction, most 2D ray casting algorithms do not provide exactly consistent results for every query. However, on average error between results from each of the methods discussed is small.

\subsection{Bresenham’s Line}

First, confirm that you have the correct template for your paper size. This template has been tailored for output on the US-letter paper size. 
It may be used for A4 paper size if the paper size setting is suitably modified.

\subsection{Ray Marching Distance Transforms}

Similar to Bresenham’s Line algorithm, ray marching [citation below] checks points along the line radiating outwards from the query point until an obstacle is encountered. The primary difference is that ray marching makes larger steps along the query ray, avoiding unnecessary memory reads. Beginning at the query point (x,y) the ray marching algorithm proceeds in the theta direction, stepping along the line by the minimum distance between each query point and the nearest obstacle. The algorithm terminates when the query point falls on an obstacle in the map. A precomputed euclidean distance transform of the map provides the distance between each query point and the nearest obstacle in any direction.
Ray marching is on average much faster than Bresenham’s line, but edge cases exist in which the performance is equivalent. For example, a query ray which travels closely parallel to a wall will make small steps between each query point for the whole length of the ray. As a result, the performance of ray marching has a long tail distribution which can be problematic for near real time algorithms. Though highly scene dependent, on average ray marching runtime is roughly logarithmic in distance between obstacles, though in the worst case it is linear in map size.

[FIGURE HERE]

\subsection{Lookup Table}

As previously described, a common acceleration technique for two dimensional ray casting is to simply precompute the ray distances for every state in a discrete grid and store the results in a three dimensional LUT. Any ray casting technique may be used to compute the table. In the authors’ implementation ray casting is used to populate the LUT since it has low initialization time and is significantly faster than Bresenham’s Line. Theoretically this approach has constant query runtime, though actual performance is dependent on access patterns as CPU caching effects are significant in practice.

State space discretization implies approximate results, since intermediate states must be rounded onto the discrete grid. While the effect is small for x and y, it can be significant for theta. As the ray moves further away from the query point along the theta direction, angular discretization error accumulates. For queries $(x,y,\theta)$ discretized into $(x_d,y_d,\theta_d)$, the distance between the end of the ray $(x_d,y_d,Θ_d)$ and its projection onto the line implied by $(x,y,\theta)$ becomes large as the length of the ray increases. Although the error induced by discretization may be unacceptable for computer graphics applications, it is generally acceptable for probabilistic algorithms such as the recursive Bayes filter since error is expected and directly modeled.

One method to improve accuracy of queries which fall between discrete states is to ray cast the neighboring discrete states and interpolate between them, rather than simply rounding to the nearest discrete state. Since we are evaluating the performance of these methods for use in a particle filter, we do not perform interpolation since the increased computation reduces the number of particles that may be maintained in real time and therefore any increase in accuracy with respect to the stored map is effectively negated.

\section{APPROACH}

Although the three dimensional table used to store precomputed ray cast solutions for a discrete state space is inherently large, it is highly compressible. This is most apparent in the cardinal directions, in which adjacent values along a particular dimension of the table increase by exactly one unit of distance for unobstructed positions. Our data structure is designed to compress this redundancy while still allowing for fast queries in near constant time. We accomplish this though through what we refer to as a compressed directional distance transform (CDDT) described here.

The euclidean distance transform (EDT) of an occupancy grid map stores the distance to the nearest obstacle in any direction for each point $(x,y)$ in the map. In contrast, a directional distance transform (DDT) stores the distance to the nearest obstacle in a particular direction for each point $(x,y)$ in the map. The key difference between a single Θ slice of the LUT and a directional distance transform is the way in which it is computed. To compute the LUT slice, ray casting is performed in the $\theta_d$ direction from every $(x_d,y_d)$ in the map using another algorithm such as Bresenham’s line. In contrast, to compute the DDT, the obstacles in the map are rotated about the center of the map by $-\theta_d$ and the values of each column in the DDT may be directly computed as a sawtooth function which goes to zero whenever an obstacle is encountered in that column. Thus, each column in the DDT may be characterized as a sawtooth function as is illustrated by figure [DDT response] below. While it is true that the rotation of the occupancy grid introduces small errors as compared to ray casting for each cell, as previously discussed these small errors do not significantly impact particle filter localization performance due to the use of a probabilistic error model. In practice, errors induced by map rotation are dominated by those inherently present in distance sensors.

[DDT GRAPHICS HERE]

The conversion from a DDT to a CDDT is performed by converting the sawtooth response of each column into a set of points in which the sawtooth wave goes to zero. This conversion is lossless in the sense that each element of the table may be recovered as a function of the nearest zero point in a particular direction. If the set of zero points is sorted, then the magnitude of the saw tooth wave at any point in the original DDT may be quickly discovered by performing binary search for the nearest zero point in that column and finding the distance between the query point and that zero point. This losslessness and fast indexing allows for ray cast queries to be quickly resolved without storing the entire DDT, but rather by storing the sorted lists of zero points.
For sake of efficiency, the full DDT need not be computed, but rather the CDDT may be built directly by projecting each obstacle into the coordinate space of each theta slice and storing a list of y-coordinates for each discrete x-coordinate. Once all scene geometry is projected into the table, each list of zero points is sorted to facilitate indexing.

Once the CDDT is constructed, querying the data structure to resolve ray casts involves

\begin{enumerate}
    \item Projecting the query point into the coordinate space of the slice which corresponds to the discretized theta
    \item Finding the nearest zero point which has a y-coordinate greater than or equal to the query point in the projected coordinate space
    \item Return the distance between the query point and the nearest zero point in the projected coordinate space, which is simply the difference between y-coordinates
\end{enumerate}

Binary search is theoretically optimal for finding zero points near to query states, in our implementation we conditionally use linear search when the number of zero points is below a certain constant (we use 64) as it has a superior memory access pattern in such cases.
While conceptually simple, implementing the CDDT data structure construction and traversal routines requires careful consideration in order to capture all edge cases and to minimize unnecessary computation. In this sense, it is more complex to implement than the alternatives considered, however there are many opportunities for optimization which yield real-world speed up. To ease this burden, we provide our implementation as well as Python wrappers with an Apache 2.0 license.

\subsection{Rotational Symmetry}

A side effect of extracting the zero points from each column of the DDT is the introduction of rotational symmetry. For any given theta, the zero points accumulated by projecting scene geometry into the requisite coordinate space are mirror images of those accumulated by projecting scene geometry into the theta+pi direction. Therefore, one need only compute the CDDT for the range of discrete thetas from $[0,\pi)$ and the DDT for the range $[0,2\pi)$ may be inferred, resulting in factor of two further reduction in memory usage.
In addition to the aforementioned reduction in memory usage, rotational symmetry may be exploited in scenarios where ray casts are performed radially around a single point. While traversing the data structure to resolve a ray cast in a particular $(x,y,\theta)$, the ray cast for $(x,y,\theta+\pi)$ may be resolved with a single additional memory read. Once the index i of the nearest zero point in the theta direction $z_{\theta}$ is discovered, the index of the nearest zero point in the opposite direction is simply i-1. For example, in robots with laser scanners sweeping angles larger than 180°, this symmetry can be used to reduce the number of data structure traversals required to compute the sensor model by up to a factor of two.

\subsection{Incremental Update}

While the authors did not implement the necessary changes, we conjecture that the CDDT algorithm could be modified to allow incremental updates to the map. In contrast, any change to the map would introduce inconsistencies in a precomputed LUT requiring a full re-computation of the data structure. Ray casting every state in a three dimensional grid is highly computational, ruling out the possibility of using a precomputed LUT with dynamic maps such as in SLAM algorithms. Similarly, ray marching operates on a euclidean distance transform which cannot be easily updated in an incremental fashion. While Bresenham’s line algorithm is the slowest option evaluated in terms of ray cast performance, it has the benefit that it operates directly on the occupancy grid map and can therefore be used with dynamic maps. For this reason, Bresenham’s line is one of the most commonly used algorithms for ray casting in occupancy grids.

To allow incremental update of the CDDT, the sorted vector data structure used in our implementation should be replaced with an alternative which allows for $O(logn)$ insert and delete, such as a randomized skip list. To aid in node deletion, pointers could be maintained from each grid element to the associated leaf node in the randomized skip lists with a constant factor increase in memory usage.

In this case, the cost of inserting or deleting an occupied grid cell would be $O(theta\_discretization*\log(number of lut bin elements))$. The number of elements in each lut bin tends to be small and is bounded, so this becomes $O(theta\_discretization)$ which is likely not prohibitive for real-time performance.

Skip list traversal is likely to be less efficient than binary search on sorted vectors due to worse cache characteristics. However, the authors expect it to still be significantly faster than performing Bresenham’s line, and therefore recommend that this modification be the subject of future research.

\subsection{Pruning}

By removing entries in the CDDT which can never possibly result in a ray collision, it is possible to further reduce the memory footprint of the data structure. Consider a 3x3 block of obstacles in an otherwise empty map. The center obstacle will never be the nearest neighbor in any ray casting query, because any such query would first intersect with one of the 8 surrounding obstacles. To exploit this, one can take the morphological edge map of the occupancy grid map prior to CDDT generation without loss of generality. This can result in significant memory usage and data structure construction time reductions in dense maps where the number of obstacles adjacent to empty cells is small compared to the total number of obstacles. Prior to performing ray casting on the CDDT data structure, we check the occupancy grid to ensure that the query point itself is not an obstacle and return a distance of zero if so in order to prevent incorrect results when ray casting from the middle of obstacles which may be removed by the morphological operation.

Additionally, consider a line of obstacles aligned along the Y-axis. Every element in this line will be projected into a single list of zero points for the theta=0 slice of the CDDT. However, the middle elements of the line will never result in a collision. Any ray cast from adjacent states in the $\theta=0$ direction will return early in the occupancy grid check, and any ray cast from outside the wall will collide with one of the edge-most obstacles in that line. Therefore in the theta=0 slice, one may discard the zero points corresponding to the middle elements without introducing error. This form of optimization is simple to compute in the cardinal directions, but non-trivial for arbitrary theta not aligned with an axis. Rather than attempting to analytically determine which obstacles may be discarded, it is simpler to prune the data structure by ray casting from every possible state (as in the LUT data structure) and discard any zero point which is unused.

Pruning does also increase precomputation time rendering it incompatible with incremental update, however the reduction of memory usage is worthwhile for static maps. In addition to memory reduction, we find that pruning slightly improves runtime performance, likely as a result of improved caching characteristics.

\section{EXPERIMENTS}

We have implemented the proposed algorithm in C++, as well as Bresenham’s Line, ray marching, and the LUT approach for comparison. Our source code is available for use and analysis [TODO add a link to github], and Python wrappers are also provided for easier usage.

\subsection{Synthetic Benchmarks}

We evaluate algorithm performance in two synthetic benchmarks, using two different maps. The first benchmark computes a ray cast for each point in a uniformly spaced grid over the three dimensional state space. The second benchmark performs a large number of ray casts for randomly generated states.

[INSERT GRAPHIC HERE]

\subsection{Particle Filter Experiments}

To simulate real world operating conditions, we have implemented a particle filter localization algorithm using a beam mode sensor model. We provide statistics about the ray cast performance of each algorithm while being used to compute the sensor model, as well as end to end statistics on the number of particles able to be maintained in real-time. While care was taken in implementing the particle filter, there are undoubtedly many optimizations which could further improve end-to-end performance. Thus, these benchmarks should not be considered an upper bound on performance and should be interpreted comparatively.

Our sensor model is designed for the Hokuyo UST-10LX lidar scanner used aboard the RACECAR platform [cite this?], which features a 270° field of view. Since this FOV is in excess of 180° we exploit radial symmetry discussed in [radial symmetry section] to simultaneously ray cast in the theta and $\theta+180°$ direction when possible. As demonstrated in the below figure, this optimization reduces the number of data structure traversals required by a third as compared to the other methods evaluated, while still resolving the same number of ray casts. As is standard in such applications, we downsample the resolution of the laser scanner to reduce the number of ray casts per sensor model evaluations and to make the probability distribution over the state space less peaked.

[INSERT GRAPHIC HERE]

\section{CONCLUSIONS}

This work demonstrates that the proposed CDDT algorithm may be used in mobile robotics to accelerate sensor model computation when localizing in a two dimensional occupancy grid map. While the precomputed LUT approach is generally 1.5 to 4 times faster than the proposed algorithm, the memory footprint of the proposed data structure is significantly smaller which may be desirable for resource constrained systems. Other than the LUT approach, CDDT is significantly faster than the other methods evaluated, typically ranging from a factor of 2 to 20.

In future work we aim to further develop our approach to allow for incremental map updates without requiring a full reconstruction of the underlying acceleration data structure, a feature which currently only the slowest evaluated method supports.

\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{APPENDIX}

Appendixes should appear before the acknowledgment.

\section*{ACKNOWLEDGMENT}

Is this part necessary?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.

\begin{thebibliography}{99}

\bibitem{c1} G. O. Young, ÒSynthetic structure of industrial plastics (Book style with paper title and editor),Ó 	in Plastics, 2nd ed. vol. 3, J. Peters, Ed.  New York: McGraw-Hill, 1964, pp. 15Ð64.
\bibitem{c2} W.-K. Chen, Linear Networks and Systems (Book style).	Belmont, CA: Wadsworth, 1993, pp. 123Ð135.
\bibitem{c3} H. Poor, An Introduction to Signal Detection and Estimation.   New York: Springer-Verlag, 1985, ch. 4.
\bibitem{c4} B. Smith, ÒAn approach to graphs of linear forms (Unpublished work style),Ó unpublished.
\bibitem{c5} E. H. Miller, ÒA note on reflector arrays (Periodical styleÑAccepted for publication),Ó IEEE Trans. Antennas Propagat., to be publised.
\bibitem{c6} J. Wang, ÒFundamentals of erbium-doped fiber amplifiers arrays (Periodical styleÑSubmitted for publication),Ó IEEE J. Quantum Electron., submitted for publication.
\bibitem{c7} C. J. Kaufman, Rocky Mountain Research Lab., Boulder, CO, private communication, May 1995.
\bibitem{c8} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ÒElectron spectroscopy studies on magneto-optical media and plastic substrate interfaces(Translation Journals style),Ó IEEE Transl. J. Magn.Jpn., vol. 2, Aug. 1987, pp. 740Ð741 [Dig. 9th Annu. Conf. Magnetics Japan, 1982, p. 301].
\bibitem{c9} M. Young, The Techincal Writers Handbook.  Mill Valley, CA: University Science, 1989.
\bibitem{c10} J. U. Duncombe, ÒInfrared navigationÑPart I: An assessment of feasibility (Periodical style),Ó IEEE Trans. Electron Devices, vol. ED-11, pp. 34Ð39, Jan. 1959.
\bibitem{c11} S. Chen, B. Mulgrew, and P. M. Grant, ÒA clustering technique for digital communications channel equalization using radial basis function networks,Ó IEEE Trans. Neural Networks, vol. 4, pp. 570Ð578, July 1993.
\bibitem{c12} R. W. Lucky, ÒAutomatic equalization for digital communication,Ó Bell Syst. Tech. J., vol. 44, no. 4, pp. 547Ð588, Apr. 1965.
\bibitem{c13} S. P. Bingulac, ÒOn the compatibility of adaptive controllers (Published Conference Proceedings style),Ó in Proc. 4th Annu. Allerton Conf. Circuits and Systems Theory, New York, 1994, pp. 8Ð16.
\bibitem{c14} G. R. Faulhaber, ÒDesign of service systems with priority reservation,Ó in Conf. Rec. 1995 IEEE Int. Conf. Communications, pp. 3Ð8.
\bibitem{c15} W. D. Doyle, ÒMagnetization reversal in films with biaxial anisotropy,Ó in 1987 Proc. INTERMAG Conf., pp. 2.2-1Ð2.2-6.
\bibitem{c16} G. W. Juette and L. E. Zeffanella, ÒRadio noise currents n short sections on bundle conductors (Presented Conference Paper style),Ó presented at the IEEE Summer power Meeting, Dallas, TX, June 22Ð27, 1990, Paper 90 SM 690-0 PWRS.
\bibitem{c17} J. G. Kreifeldt, ÒAn analysis of surface-detected EMG as an amplitude-modulated noise,Ó presented at the 1989 Int. Conf. Medicine and Biological Engineering, Chicago, IL.
\bibitem{c18} J. Williams, ÒNarrow-band analyzer (Thesis or Dissertation style),Ó Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, 1993. 
\bibitem{c19} N. Kawasaki, ÒParametric study of thermal and chemical nonequilibrium nozzle flow,Ó M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
\bibitem{c20} J. P. Wilkinson, ÒNonlinear resonant circuit devices (Patent style),Ó U.S. Patent 3 624 12, July 16, 1990.

\end{thebibliography}

\end{document}

### 2017-02-24 generate_serialized.sh

@@ -1,0 +1,4 @@
+./bin/range_lib --method="pcddt" --map_path=../maps/synthetic.map.png --cddt_save_path=../tmp/synthetic/pcddt.json
+./bin/range_lib --method="cddt" --map_path=../maps/synthetic.map.png --cddt_save_path=../tmp/synthetic/cddt.json
+./bin/range_lib --method="pcddt" --map_path=../maps/basement_hallways_5cm.png --cddt_save_path=../tmp/basement/pcddt.json
+./bin/range_lib --method="cddt" --map_path=../maps/basement_hallways_5cm.png --cddt_save_path=../tmp/basement/cddt.json

### 2017-02-24 includes/RangeLib.h

@@ -88,6 +88,18 @@ Useful Links: https://github.com/MRPT/mrpt/blob/4137046479222f3a71b5c00aee1d5fa8
 #define ANIL 
 #endif
 
+// these defines are for yaml serialization
+
+#define T1 "  "
+#define T2 T1 T1
+#define T3 T1 T1 T1
+#define T4 T2 T2
+
+#define J1 "  "
+#define J2 J1 J1
+#define J3 J1 J1 J1
+#define J4 J2 J2
+
 namespace ranges {
 	struct OMap
 	{
@@ -1176,6 +1188,90 @@ namespace ranges {
 					return std::make_pair(val - lut_space_x, lut_space_x - (*lut_bin)[inverse_index]);
 				}
 			}
+		}
+
+		void serializeYaml(std::stringstream* ss) {
+			// (*ss) << std::fixed;
+    		(*ss) << std::setprecision(7);
+
+			(*ss) << "cddt:" << std::endl;
+			(*ss) << T1 << "theta_discretization: " << theta_discretization << std::endl;
+			(*ss) << T1 << "lut_translations: ";
+			utils::serialize(lut_translations, ss);
+			(*ss) << std::endl;
+			(*ss) << T1 << "max_range: " << max_range << std::endl;
+			(*ss) << T1 << "map: " << std::endl;
+			(*ss) << T2 << "# note: map data is width and then height (width is number of rows) transposed from expectation:"  << std::endl;
+			(*ss) << T2 << "path: " << map.fn << std::endl;
+			(*ss) << T2 << "width: " << map.width << std::endl;
+			(*ss) << T2 << "height: " << map.height << std::endl;
+			(*ss) << T2 << "data: " << std::endl;
+			for (int i = 0; i < map.width; ++i) {
+				(*ss) << T3 << "- "; utils::serialize(map.grid[i], ss);(*ss) << std::endl;
+			}
+			(*ss) << T1 << "compressed_lut: " << std::endl;
+			for (int i = 0; i < compressed_lut.size(); ++i) {
+				#if _USE_CACHED_CONSTANTS
+				float angle = i * M_2PI_div_theta_discretization;
+				#else
+				float angle = M_2PI * i / theta_discretization;
+				#endif
+
+				(*ss) << T2 << "- slice: " << std::endl;
+				(*ss) << T3 << "theta: " << angle << std::endl;
+				(*ss) << T3 << "zeros: " << std::endl;
+
[…]

### 2017-02-24 includes/RangeUtils.h

@@ -206,6 +206,39 @@ namespace utils {
 		// std::vector<float> cache;
 		float cache[10000];
 	};
+
+	void serialize(std::vector<bool> &vals,std::stringstream* ss) {
+		if (vals.size() == 0) {
+			(*ss) << "[]";
+			return; 
+		}
+		(*ss) << "[" << vals[0];
+		for (int i = 1; i < vals.size(); ++i) {
+			(*ss) << "," << vals[i];
+		}
+		(*ss) << "]";
+	}
+
+	void serialize(std::vector<float> &vals,std::stringstream* ss) {
+		if (vals.size() == 0) {
+			(*ss) << "[]";
+			return; 
+		}
+		(*ss) << "[" << vals[0];
+		for (int i = 1; i < vals.size(); ++i) {
+			(*ss) << "," << vals[i];
+		}
+		(*ss) << "]";
+	}
+
+	std::string serialize(std::vector<float> &vals) {
+		std::stringstream ss;
+		serialize(vals,&ss);
+		return ss.str();
+	}
+
+
+
 } // namespace utils
 
 #endif	/* _RANGE_UTILS_H_INCLUDED_ */

### 2017-02-24 main.cpp

@@ -29,6 +29,9 @@ DEFINE_string(method, "RayMarching",
 DEFINE_string(map_path, "BASEMENT_MAP",
 	"Path to map image, relative to current directory");
 
+DEFINE_string(cddt_save_path, "",
+	"Path to serialize CDDT data structure to.");
+
 #define MAX_DISTANCE 500
 #define NUM_RAYS 500
 #define THETA_DISC 108
@@ -69,15 +72,16 @@ int main(int argc, char *argv[])
 	OMap map = OMap(1,1);
 	if (FLAGS_map_path == "BASEMENT_MAP") {
 		#ifdef BASEPATH
-		#pragma GCC diagnostic push
-		#pragma GCC diagnostic ignored "-Wunused-result"
-		(volatile void) chdir(BASEPATH);
-		#pragma GCC diagnostic pop
+		std::cout << "...Loading map" << std::endl;
+		map = OMap(BASEPATH "/maps/basement_hallways_5cm.png");
+		// #pragma GCC diagnostic push
+		// #pragma GCC diagnostic ignored "-Wunused-result"
+		// (volatile void) chdir(BASEPATH);
+		// #pragma GCC diagnostic pop
 		#else
 		std::cout << "BASEPATH not defined, map paths may not resolve correctly." << std::endl;
 		#endif
-		std::cout << "...Loading map" << std::endl;
-		map = OMap("./maps/basement_hallways_5cm.png");
+		
 	} else {
 		map = OMap(FLAGS_map_path);
 	}
@@ -157,6 +161,13 @@ int main(int argc, char *argv[])
 			#ifdef VERBOSE_LOG_PATH
 			save_log(log, VERBOSE_LOG_PATH "/pcddt.csv");
 			#endif
+		}
+
+		if (!FLAGS_cddt_save_path.empty()) {\
+			std::cout << "...saving CDDT to:" << FLAGS_cddt_save_path<< std::endl;
+			std::stringstream cddt_serialized;
+			rc.serializeJson(&cddt_serialized);
+			save_log(cddt_serialized, FLAGS_cddt_save_path.c_str());
 		}
 	}
 

### 2017-02-24 visualize.py

@@ -1,0 +1,237 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import yaml
+from yaml import CLoader as Loader, CDumper as Dumper
+import ujson
+import itertools
+import argparse
+
+# dump = yaml.dump(dummy_data, fh, encoding='utf-8', default_flow_style=False, Dumper=Dumper)
+# data = yaml.load(fh, Loader=Loader)
+
+parser = argparse.ArgumentParser()
+parser.add_argument('--path', help='Path to serialized json CDDT data structure')
+
+class Map(object):
+    """ Map saved in a serialized CDDT """
+    def __init__(self, data):
+        print "...loading map"
+        self.path = data["path"]
+        self.width = data["width"]
+        self.height = data["height"]
+        self.data = np.array(data["data"]).transpose()
+
+    def visualize(self):
+        plt.imshow(-1*self.data, cmap="gray")
+        plt.show()
+
+class CDDTSlice(object):
+    """ Contains a single slice of CDDT corresponding to a single theta value"""
+    def __init__(self, data):
+        # print "...loading slice"
+        self.theta = data["theta"]
+        self.zeros = data["zeros"]
+
+    def num_zeros(self):
+        return [len(lut_bin) for lut_bin in self.zeros]
+
+    def ddt_dims(self):
+        non_empty_zeros = filter(lambda x: len(x) > 0, self.zeros)
+        min_zero = min(map(min, non_empty_zeros))
+        max_zero = max(map(max, non_empty_zeros))
+        return [int(np.ceil(max_zero - min_zero))+1,len(self.zeros)]
+
+    def make_ddt(self, saw_tooth=True):
+        non_empty_zeros = filter(lambda x: len(x) > 0, self.zeros)
+        if len(non_empty_zeros) == 0:
+            print "Empty slice, nothing to visualize"
+            return
+
+        # print map(min, self.zeros)
+        min_zero = min(map(min, non_empty_zeros))
+        max_zero = max(map(max, non_empty_zeros))
+        height = int(np.ceil(max_zero - min_zero))+1
+        ddt = np.zeros((height,len(self.zeros)))
+
+        for x in xrange(len(self.zeros)):
+            for zp in self.zeros[x]:
+                y = int(zp - min_zero)
+                ddt[y,x] = 1
+
+        if saw_tooth:
[…]

### 2017-02-25 paper/main.tex
paper changes

@@
 % The following packages can be found on http:\\www.ctan.org
-%\usepackage{graphics} % for pdf, bitmapped graphics files
+\usepackage{graphicx}
+\usepackage{tabu}
+% \usepackage{graphics} % for pdf, bitmapped graphics files
 %\usepackage{epsfig} % for postscript graphics files
@@
 %\usepackage{amssymb}  % assumes amsmath package installed
+
+\newcommand{\img}[1]{\begin{center}\includegraphics[width=\columnwidth]{{#1}}\end{center}}
+
+\newcommand{\imgFrame}[1]{\begin{center}\frame{\includegraphics[width=\columnwidth]{{#1}}}\end{center}}
+
+% \begin{center}\frame{\includegraphics[width=\columnwidth]{synthetic.png}}\end{center}
 
@@
 
-The ray cast operation is an important sub-problem for robot localization using distance sensors to detect the environment.
+The ray cast operation is an important sub-problem for robot localization \cite{localization} using distance sensors to detect the environment.
 In particle filter localization, particles are re-weighted according to a Bayesian update equation which relies on ray casting to determine the ground truth distance between the hypothesis pose and obstacles in a map.
@@
 To combat this challenge, Fox et al.
-[1999] suggest the use of a large three-dimensional lookup table (LUT) to store expected ranges for each discrete (x,y,Θ) state.
+\cite{localization} suggest the use of a large three-dimensional lookup table (LUT) to store expected ranges for each discrete $(x,y,\theta)$ state.
 While this is simple to implement and does result in large speed improvements as compared to ray casting, it can be prohibitively memory intensive for large maps and/or resource constrained systems.
@@
 Many robotic applications require three dimensional maps, however that is considered outside the scope of this work since many data structures exist for accelerating 3D ray casting, such as the bounded volume hierarchy and related variants.
-The five dimensional table required to precompute expected ranges for a discrete state space (x,y,z,Θ,Φ) would be prohibitively expensive to both compute and store, so 3D acceleration data structures are exclusively used when 3D ray casting is required.
+The five dimensional table required to precompute expected ranges for a discrete state space $(x,y,z,\theta,\phi)$ would be prohibitively expensive to both compute and store, so 3D acceleration data structures are exclusively used when 3D ray casting is required.
 Though our approach does extend to higher dimensionality and would be smaller than the analogous five dimensional table, it would still be prohibitively expensive in greater than two dimensions, thus we do not consider it here.
@@
 
-\subsection{Bresenham’s Line}
+\subsection{Bresenham's Line}
 
-First, confirm that you have the correct template for your paper size. This template has been tailored for output on the US-letter paper size. 
-It may be used for A4 paper size if the paper size setting is suitably modified.
+Bresenham's line algorithm \cite{bresenham} incrementally determines the set of pixels that approximate the trajectory of query ray starting from the query point $(x,y)$ and progressing in the $\theta$ direction.
+The algorithm terminates once the nearest occupied pixel is discovered, and the euclidean distance between that occupied pixel and the query $(x,y)$ is reported.
+This algorithm is widely implemented in particle filters due to its simplicity and ability to operate on a dynamic map.
+The primary disadvantage is that it is slow, potentially requiring hundreds of memory accesses for a single ray cast.
+While actual performance is highly environment dependent, Bresenham’s Line algorithm is linear in map size in the worst case.
 
@@
 
-Similar to Bresenham’s Line algorithm, ray marching [citation below] checks points along the line radiating outwards from the query point until an obstacle is encountered.
+Similar to Bresenham's Line algorithm, ray marching \cite{raymarching} checks points along the line radiating outwards from the query point until an obstacle is encountered.
 The primary difference is that ray marching makes larger steps along the query ray, avoiding unnecessary memory reads.
-Beginning at the query point (x,y) the ray marching algorithm proceeds in the theta direction, stepping along the line by the minimum distance between each query point and the nearest obstacle.
+Beginning at the query point $(x,y)$ the ray marching algorithm proceeds in the theta direction, stepping along the line by the minimum distance between each query point and the nearest obstacle.
 The algorithm terminates when the query point falls on an obstacle in the map.
 A precomputed euclidean distance transform of the map provides the distance between each query point and the nearest obstacle in any direction.
+
+\begin{figure}[h]
+\img{spheretrace.jpg}
+\caption{Visualization of ray marching starting at $p_0$ towards $p_4$.
+Green circle around each query point represents the distance to the nearest obstacle from that point.
+Blue dots represent the query points, labeled in order of execution.
+From \cite{sphere:source}.}
+\label{spheres}
+\end{figure}
+
 Ray marching is on average much faster than Bresenham’s line, but edge cases exist in which the performance is equivalent.
@@
 
-[FIGURE HERE]
+% \begin{figure}[thpb]
+%       \centering
+%       \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
+% }}
+%       %\includegraphics[scale=1.0]{figurefile}
+%       \caption{Inductance of oscillation winding on amorphous
+%       magnetic core versus DC bias magnetic field}
+%       \label{figurelabel}
+%   \end{figure}
 
@@
 
+\begin{figure}[h]
+\img{basement_bw_cropped.png}
+\caption{Occupancy grid map of the Stata basement loop. 1200x1200}
+\label{stata}
+\end{figure}
+
 The euclidean distance transform (EDT) of an occupancy grid map stores the distance to the nearest obstacle in any direction for each point $(x,y)$ in the map.
@@
 In contrast, to compute the DDT, the obstacles in the map are rotated about the center of the map by $-\theta_d$ and the values of each column in the DDT may be directly computed as a sawtooth function which goes to zero whenever an obstacle is encountered in that column.
-Thus, each column in the DDT may be characterized as a sawtooth function as is illustrated by figure [DDT response] below.
-While it is true that the rotation of the occupancy grid introduces small errors as compared to ray casting for each cell, as previously discussed these small errors do not significantly impact particle filter localization performance due to the use of a probabilistic error model.
+Thus, each column in the DDT may be characterized as a sawtooth function with uneven periodicity.
+While it is true that the rotation of the occupancy grid introduces small errors as compared to ray casting for each cell, as previously discussed small errors do not significantly impact particle filter localization performance due to the use of a probabilistic error model.
 In practice, errors induced by map rotation are dominated by those inherently present in distance sensors.
 
-[DDT GRAPHICS HERE]
+\begin{figure}[h]
+\img{slice0_cddt_basement.png}
+\caption{A single slice of the DDT generated from the CDDT data structure for $\theta=0$.
+The value of each pixel represents the distance to the nearest obstacle in the upwards direction with respect to the image above.
+Each column may be thought of as a sawtooth wave, and the points where the wave is zero are projected into the CDDT bin for that column for later indexing.
+The blue plot indicates the number of zero points projected into each CDDT bin.
+The compression factor of 267 is computed as two times the size of the corresponding LUT slice divided by the size of this slice in the CDDT.
+The factor of two is due to the use of rotational symmetry to reduce the number of slices maintained in the CDDT with respect to the LUT.}
+\label{cddt:slice0}
+\end{figure}
+
+\begin{figure}[h]
+\img{slice41_cddt_basement.png}
+\caption{Another slice of the DDT generated from the CDDT data structure for $\theta=2.385$.}
+\label{cddt:slice41}
+\end{figure}
 
@@
 Therefore, one need only compute the CDDT for the range of discrete thetas from $[0,\pi)$ and the DDT for the range $[0,2\pi)$ may be inferred, resulting in factor of two further reduction in memory usage.
+
 In addition to the aforementioned reduction in memory usage, rotational symmetry may be exploited in scenarios where ray casts are performed radially around a single point.
@@
 
+\begin{figure}[h]
+\img{slice0_pcddt_basement.png}
+\caption{The same slice of the DDT as \ref{cddt:slice0}.
+Notice that the number of elements projected into each CDDT bin is less peaky since the pruning operation removes most elements of long lines of obstacles.
+Also note the higher compression factor of 415 for this slice, increased from 267 in the unpruned CDDT.}
+\label{pcddt:slice0}
+\end{figure}
+
 Additionally, consider a line of obstacles aligned along the Y-axis.
@@
 
+\begin{figure}[h]
+\img{synthetic_cropped.png}
+\caption{Occupancy grid map for the synthetic map example.}
+\label{violin:basement:grid}
+\end{figure}
+
 We have implemented the proposed algorithm in C++, as well as Bresenham’s Line, ray marching, and the LUT approach for comparison.
 Our source code is available for use and analysis [TODO add a link to github], and Python wrappers are also provided for easier usage.
+All benchmarks were performed on a desktop computer on a Intel Core i5-4590 CPU @ 3.30GHz with 16GB DDR3 ram @ 1333MHz, running Ubuntu 14.04.
 
@@
 
+\begin{figure}[h]
+\img{grid_trial2_basement_cropped.png}
+\caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method.
+Basement map.
+X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
+\label{violin:basement:grid}
+\end{figure}
+
+\begin{figure}[h]
+\img{grid_trial2_synthetic_cropped.png}
+\caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method.
+Synthetic map.
+X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
+\label{violin:basement:grid}
+\end{figure}
+
+\begin{figure}[h]
+\img{random_trial2_basement_cropped.png}
+\caption{Violin plots for a large number of randomly sampled queries. Basement map. The random access pattern degrades the caching performance variably for each method.}
+\label{violin:basement:random}
+\end{figure}
+
+\begin{figure}[h]
+\img{random_trial2_synthetic_cropped.png}
+\caption{Violin plots for a large number of randomly sampled queries. Synthetic map. The random access pattern degrades the caching performance variably for each method.}
+\label{violin:basement:random}
+\end{figure}
+
 We evaluate algorithm performance in two synthetic benchmarks, using two different maps.
+The so called Synthetic map was created with photoshop, whereas the basement map was created via a SLAM algorithm on the RACECAR platform [cite this?].
 The first benchmark computes a ray cast for each point in a uniformly spaced grid over the three dimensional state space.
@@
 
-[INSERT GRAPHIC HERE]
+\begin{figure}[h]
+\begin{center}
+\begin{tabular}{ | m{8em} | m{2.3cm}| m{1.5cm} | } 
+\hline
+ \multicolumn{3}{|c|}{Basement Map, $\theta$ discretization: 108} \\
+ \hline
+ Method & Memory Usage & Init. Time \\
+ \hline
+ Bresenham's Line & 1.37 MB & 0.006 sec  \\
+ Ray Marching & 5.49 MB & 0.16 sec  \\
+ CDDT & 6.34 MB & 0.067 sec  \\
+ PCDDT & 4.07 MB & 2.2 sec  \\
+ Lookup Table & 296.63 MB & 15.3 sec  \\
+\hline
+\end{tabular}
+\end{center}
+\caption{The construction time and memory footprint of each method for the Basement map.}
+\label{table:basement:init}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\begin{tabular}{ | m{8em} | m{2.3cm}| m{1.5cm} | } 
+\hline
+ \multicolumn{3}{|c|}{Synthetic Map, $\theta$ discretization: 108} \\
+ \hline
+ Method & Memory Usage & Init. Time \\
+ \hline
+ Bresenham's Line & 1 MB & 0.004 sec  \\
+ Ray Marching  & 4 MB & 0.13 sec  \\
+ CDDT & 2.71 MB & 0.03 sec  \\
+ PCDDT & 1.66 MB & 0.74 sec  \\
+ Lookup Table & 216 MB & 9.1 sec  \\
+\hline
+\end{tabular}
+\end{center}
+\caption{The construction time and memory footprint of each method for the Basement map.}
+\label{table:basement:init}
+\end{figure}
 
@@
 
-[INSERT GRAPHIC HERE]
+\begin{figure}[h]
+\img{particle_trial2_basement_cropped.png}
+\caption{Violin plots for ray casting completion times during the execution of a particle filter.
+Basement map.
+Ray casting is primarily done inside of hallways, reducing the mean ray length and thereby improving Bresenham's Line performance.
+CDDT and PCDDT exploiting radial symmetry while resolving the sensor model.}
+\label{violin:basement:random}
+\end{figure}
 
@@
 
-Appendixes should appear before the acknowledgment.
+\begin{figure}[h]
+\begin{center}
+\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
+\hline
+\multicolumn{5}{|c|}{Basement Map, Random Sampling Ray Cast} \\
+\hline
+Method & Mean & Median & IQR & Speedup \\
+\hline
+BL & 9.66e-07 & 1.05e-06 & 1.08e-06 & 1 \\
+RM & 2.12e-07 & 1.68e-07 & 1.64e-07 & 4.56 \\
+CDDT & 1.58e-07 & 1.49e-07 & 9.1e-08 & 6.13 \\
+PCDDT & 1.41e-07 & 1.32e-07 & 6.5e-08 & 6.83 \\
+LUT & 1.89e-07 & 1.7e-07 & 2.1e-08 & 5.10 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{The construction time and memory footprint of each method for the Basement map.}
+\label{table:basement:init}
+\end{figure}
 
-\section*{ACKNOWLEDGMENT}
+\begin{figure}[h]
+\begin{center}
+\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
+\hline
+\multicolumn{5}{|c|}{Basement Map, Grid Sampling Ray Cast} \\
+\hline
+Method & Mean & Median & IQR & Speedup \\
+\hline
+BL & 8.29e-07 & 9.24e-07 & 9.53e-07 & 1 \\
+RM & 1.65e-07 & 1.26e-07 & 1.34e-07 & 5.02 \\
+CDDT & 7.69e-08 & 7.2e-08 & 1.6e-08 & 10.78 \\
+PCDDT & 7.32e-08 & 7e-08 & 1.4e-08 & 11.33 \\
+LUT & 6.13e-08 & 4.3e-08 & 4.6e-08 & 13.53 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{The construction time and memory footprint of each method for the Basement map.}
+\label{table:basement:init}
+\end{figure}
 
-Is this part necessary?
+\begin{figure}[h]
+\begin{center}
+\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
+\hline
+\multicolumn{5}{|c|}{Basement Map, Particle Filter Ray Cast} \\
+\hline
+Method & Mean & Median & IQR & Speedup \\
+\hline
+BL & 3.89e-07 & 2.67e-07 & 2.31e-07 & 1 \\
+RM & 1.56e-07 & 1.18e-07 & 9.4e-08 & 2.49 \\
+CDDT & 7.67e-08 & 7.8e-08 & 5.7e-08 & 5.07 \\
+PCDDT & 7.22e-08 & 7.4e-08 & 5.4e-08 & 5.39 \\
+LUT & 4.35e-08 & 4.1e-08 & 5e-09 & 8.94 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{The construction time and memory footprint of each method for the Basement map.}
+\label{table:basement:init}
+\end{figure}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[h]
+\begin{center}
+\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
+\hline
+\multicolumn{5}{|c|}{Synthetic Map, Random Sampling Ray Cast} \\
+\hline
+Method & Mean & Median & IQR & Speedup \\
+\hline
+BL & 1.19e-06 & 1.41e-06 & 7.71e-07 & 1 \\
+RM & 1.52e-07 & 1.25e-07 & 1.05e-07 & 7.81 \\
+CDDT & 1.24e-07 & 1.05e-07 & 5.3e-08 & 9.59 \\
+PCDDT & 1.19e-07 & 1.01e-07 & 5e-08 & 10.02 \\ 
+LUT & 1.82e-07 & 1.68e-07 & 1.4e-08 & 6.55 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{The construction time and memory footprint of each method for the Basement map.}
+\label{table:basement:init}
+\end{figure}
 
-References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.
+\begin{figure}[h]
+\begin{center}
+\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
+\hline
+\multicolumn{5}{|c|}{Synthetic Map, Grid Sampling Ray Cast} \\
+\hline
+Method & Mean & Median & IQR & Speedup \\
+\hline
+BL & 1.03e-06 & 1.20e-06 & 6.79e-07 & 1 \\ 
+RM & 1.27e-07 & 1.03e-07 & 1.06e-07 & 8.06 \\
+CDDT & 7.02e-08 & 6.8e-08 & 1e-08 & 14.63 \\
+PCDDT & 6.94e-08 & 6.8e-08 & 9e-09 & 14.80 \\
+LUT & 6.33e-08 & 4.2e-08 & 4.6e-08 & 16.21 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{The construction time and memory footprint of each method for the Basement map.}
+\label{table:basement:init}
+\end{figure}
 
@@
 
-\bibitem{c1} G. O. Young, ÒSynthetic structure of industrial plastics (Book style with paper title and editor),Ó 	in Plastics, 2nd ed. vol. 3, J. Peters, Ed.  New York: McGraw-Hill, 1964, pp. 15Ð64.
-\bibitem{c2} W.-K. Chen, Linear Networks and Systems (Book style).	Belmont, CA: Wadsworth, 1993, pp. 123Ð135.
-\bibitem{c3} H. Poor, An Introduction to Signal Detection and Estimation.   New York: Springer-Verlag, 1985, ch. 4.
-\bibitem{c4} B. Smith, ÒAn approach to graphs of linear forms (Unpublished work style),Ó unpublished.
-\bibitem{c5} E. H. Miller, ÒA note on reflector arrays (Periodical styleÑAccepted for publication),Ó IEEE Trans. Antennas Propagat., to be publised.
-\bibitem{c6} J. Wang, ÒFundamentals of erbium-doped fiber amplifiers arrays (Periodical styleÑSubmitted for publication),Ó IEEE J. Quantum Electron., submitted for publication.
-\bibitem{c7} C. J. Kaufman, Rocky Mountain Research Lab., Boulder, CO, private communication, May 1995.
-\bibitem{c8} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ÒElectron spectroscopy studies on magneto-optical media and plastic substrate interfaces(Translation Journals style),Ó IEEE Transl.
-J.
-Magn.Jpn., vol.
-2, Aug.
-1987, pp.
-740Ð741 [Dig.
-9th Annu.
-Conf.
-Magnetics Japan, 1982, p.
-301].
-\bibitem{c9} M. Young, The Techincal Writers Handbook.  Mill Valley, CA: University Science, 1989.
-\bibitem{c10} J. U. Duncombe, ÒInfrared navigationÑPart I: An assessment of feasibility (Periodical style),Ó IEEE Trans. Electron Devices, vol. ED-11, pp. 34Ð39, Jan. 1959.
-\bibitem{c11} S.
-Chen, B.
-Mulgrew, and P.
-M.
-Grant, ÒA clustering technique for digital communications channel equalization using radial basis function networks,Ó IEEE Trans.
-Neural Networks, vol.
-4, pp.
-570Ð578, July 1993.
-\bibitem{c12} R. W. Lucky, ÒAutomatic equalization for digital communication,Ó Bell Syst. Tech. J., vol. 44, no. 4, pp. 547Ð588, Apr. 1965.
-\bibitem{c13} S.
-P.
-Bingulac, ÒOn the compatibility of adaptive controllers (Published Conference Proceedings style),Ó in Proc.
-4th Annu.
-Allerton Conf.
-Circuits and Systems Theory, New York, 1994, pp.
-8Ð16.
-\bibitem{c14} G. R. Faulhaber, ÒDesign of service systems with priority reservation,Ó in Conf. Rec. 1995 IEEE Int. Conf. Communications, pp. 3Ð8.
-\bibitem{c15} W. D. Doyle, ÒMagnetization reversal in films with biaxial anisotropy,Ó in 1987 Proc. INTERMAG Conf., pp. 2.2-1Ð2.2-6.
-\bibitem{c16} G.
-W.
-Juette and L.
-E.
-Zeffanella, ÒRadio noise currents n short sections on bundle conductors (Presented Conference Paper style),Ó presented at the IEEE Summer power Meeting, Dallas, TX, June 22Ð27, 1990, Paper 90 SM 690-0 PWRS.
-\bibitem{c17} J. G. Kreifeldt, ÒAn analysis of surface-detected EMG as an amplitude-modulated noise,Ó presented at the 1989 Int. Conf. Medicine and Biological Engineering, Chicago, IL.
-\bibitem{c18} J. Williams, ÒNarrow-band analyzer (Thesis or Dissertation style),Ó Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, 1993. 
-\bibitem{c19} N. Kawasaki, ÒParametric study of thermal and chemical nonequilibrium nozzle flow,Ó M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
-\bibitem{c20} J. P. Wilkinson, ÒNonlinear resonant circuit devices (Patent style),Ó U.S. Patent 3 624 12, July 16, 1990.
+\bibitem{localization} Sebastian Thrun, Dieter Fox, Wolfram Burgard and Frank Dellaert. “Robust Monte Carlo Localization for Mobile Robots.” Artificial Intelligence Journal. 2001
+
+\bibitem{bresenham} J. Bresenham, “Algorithm for Computer Control of a Digital Plotter,” IBM Systems Journal, vol. 4, no. 1, pp. 25-30, 1965
+
+\bibitem{raymarching} K. Perlin and E. M. Hoffert, “Hypertexture.” Computer Graphics, vol 23, no. 3, pp. 297-306, 1989.
+
+\bibitem{sphere:source} Pharr, M., and R.
+Fernando.
+GPU Gems 2: Programming Techniques for High-Performance Graphics and General-Purpose Computation: Addison Wesley Professional.
+2005.
+http://http.developer.nvidia.com/GPUGems2/gpugems2\_chapter08.html
 
 \end{thebibliography}
-
 \end{document}

### 2017-02-25 generate_serialized.sh

@@ -1,4 +1,11 @@
-./bin/range_lib --method="pcddt" --map_path=../maps/synthetic.map.png --cddt_save_path=../tmp/synthetic/pcddt.json
-./bin/range_lib --method="cddt" --map_path=../maps/synthetic.map.png --cddt_save_path=../tmp/synthetic/cddt.json
-./bin/range_lib --method="pcddt" --map_path=../maps/basement_hallways_5cm.png --cddt_save_path=../tmp/basement/pcddt.json
-./bin/range_lib --method="cddt" --map_path=../maps/basement_hallways_5cm.png --cddt_save_path=../tmp/basement/cddt.json
+# ./bin/range_lib --method="pcddt" --map_path=../maps/synthetic.map.png --cddt_save_path=../tmp/synthetic/pcddt.json
+# ./bin/range_lib --method="cddt" --map_path=../maps/synthetic.map.png --cddt_save_path=../tmp/synthetic/cddt.json
+# ./bin/range_lib --method="pcddt" --map_path=../maps/basement_hallways_5cm.png --cddt_save_path=../tmp/basement/pcddt.json
+# ./bin/range_lib --method="cddt" --map_path=../maps/basement_hallways_5cm.png --cddt_save_path=../tmp/basement/cddt.json
+
+
+./bin/range_lib --method="cddt" --map_path=../maps/basement_hallways_5cm.png --log_path=../tmp/basement/trial1/
+./bin/range_lib --method="cddt" --map_path=../maps/basement_hallways_5cm.png --log_path=../tmp/basement/trial2/
+./bin/range_lib --method="cddt" --map_path=../maps/basement_hallways_5cm.png --log_path=../tmp/basement/trial3/
+./bin/range_lib --method="cddt" --map_path=../maps/basement_hallways_5cm.png --log_path=../tmp/basement/trial4/
+./bin/range_lib --method="cddt" --map_path=../maps/basement_hallways_5cm.png --log_path=../tmp/basement/trial5/

### 2017-02-25 includes/RangeLib.h

@@ -284,6 +284,12 @@ namespace ranges {
 		bool error() {
 			return has_error;
 		}
+
+		// returns memory usage in bytes
+		int memory() {
+			// std::cout << sizeof(bool) * width * height << std::endl;
+			return sizeof(bool) * width * height;
+		}
 	};
 
 	struct DistanceTransform
@@ -348,6 +354,10 @@ namespace ranges {
 			//if there's an error, display it
 			if(error) std::cout << "encoder error " << error << ": "<< lodepng_error_text(error) << std::endl;
 			return error;
+		}
+
+		int memory() {
+			return width*height*sizeof(float);
 		}
 	};
 
@@ -463,7 +473,7 @@ namespace ranges {
 			return max_range; 
 		}
 
-		int memory() { return map.width*map.height; }
+		int memory() { return map.memory(); }
 	};
 
 	class RayMarching : public RangeMethod
@@ -508,7 +518,7 @@ namespace ranges {
 			return max_range; 
 		}
 
-		int memory() { return map.width*map.height*2; }
+		int memory() { return distImage.memory(); }
 	protected:
 		DistanceTransform distImage;
 		float distThreshold = 0.0;
@@ -712,7 +722,7 @@ namespace ranges {
 			return lut_size * sizeof(float);
 		}
 
-		int memory() { return lut_size(); }
+		int memory() { return lut_size()+map.memory()+lut_translations.size()*sizeof(float); }
 
 		// mark all of the LUT entries that are potentially useful
 		// remove all LUT entries that are not potentially useful
@@ -1352,9 +1362,6 @@ namespace ranges {
 						#else
 						lut_row.push_back(r);
 						#endif
-
-
-						
 					}
 					lut_slice.push_back(lut_row);
 				}
@@ -1490,6 +1497,34 @@ namespace benchmark {
 			std::cout << " -avg time per ray: " << ( t_accum / (float) num_cast) << " sec" << std::endl;
 			std::cout << " -rays cast: " << num_cast << std::endl;
 			std::cout << " -total time: " << t_accum << " sec" << std::endl;
+		}
+
+		void random_sample(int num_samples) {
+			std::default_random_engine generator;
+			generator.seed(clock());
+			std::uniform_real_distribution<float> randx = std::uniform_real_distribution<float>(1.0,map->width - 1.0);
+			std::uniform_real_distribution<float> randy = std::uniform_real_distribution<float>(1.0,map->height - 1.0);
[…]

### 2017-02-25 main.cpp

@@ -32,24 +32,29 @@ DEFINE_string(map_path, "BASEMENT_MAP",
 DEFINE_string(cddt_save_path, "",
 	"Path to serialize CDDT data structure to.");
 
+DEFINE_string(log_path, "",
+	"Where to store high fidelity logs, does not save if not specified.");
+
+DEFINE_string(which_benchmark, "random",
+	"Which benchmark to run, one of:\n  random\n  grid\n");
+
 #define MAX_DISTANCE 500
 #define NUM_RAYS 500
 #define THETA_DISC 108
 #define MB (1024.0*1024.0)
-#ifdef BASEPATH
-#define VERBOSE_LOG_PATH BASEPATH "/tmp"
-#endif
+// #ifdef BASEPATH
+// #define FLAGS_log_path BASEPATH "/tmp/basement/"
+// #endif
 
 // grid sample settings
 #define GRID_STEP 10
 #define GRID_RAYS 40
 #define GRID_SAMPLES 1
-
+#define RANDOM_SAMPLES 200000
 
 using namespace ranges;
 using namespace benchmark;
 
-#ifdef VERBOSE_LOG_PATH
 void save_log(std::stringstream &log, const char *fn) {
 	std::cout << "...saving log to: " << fn << std::endl;
 	std::ofstream file;  
@@ -57,7 +62,10 @@ void save_log(std::stringstream &log, const char *fn) {
 	file << log.str();
 	file.close();
 }
-#endif
+
+void save_log(std::stringstream &log, std::string path) {
+	save_log(log,path.c_str());
+}
 
 int main(int argc, char *argv[])
 {
@@ -86,81 +94,124 @@ int main(int argc, char *argv[])
 		map = OMap(FLAGS_map_path);
 	}
 
-	#ifdef VERBOSE_LOG_PATH
-	std::cout << "Saving logs to: " << VERBOSE_LOG_PATH "/" << std::endl;
-	std::stringstream log; 
-	#endif
-	
+	bool DO_LOG = (FLAGS_log_path != "");
+	std::stringstream tlog;
+	std::stringstream summary; 
+	if (DO_LOG) {
+		std::cout << "Saving logs to: " << FLAGS_log_path << std::endl;
+		summary << std::fixed;
+	    summary << std::setprecision(9);
+		summary << "method,construction_time,memory_bytes" << std::endl;
+	}
+
 	if (map.error()) return 1;
 
[…]

### 2017-02-25 paper/data/generate_data.sh

@@ -1,0 +1,52 @@
+mkdir -p ./basement/grid/trial1 ./basement/grid/trial2 ./basement/grid/trial3 ./basement/grid/trial4 ./basement/grid/trial5
+mkdir -p ./synthetic/grid/trial1 ./synthetic/grid/trial2 ./synthetic/grid/trial3 ./synthetic/grid/trial4 ./synthetic/grid/trial5
+mkdir -p ./basement/random/trial1 ./basement/random/trial2 ./basement/random/trial3 ./basement/random/trial4 ./basement/random/trial5
+mkdir -p ./synthetic/random/trial1 ./synthetic/random/trial2 ./synthetic/random/trial3 ./synthetic/random/trial4 ./synthetic/random/trial5
+mkdir -p ./basement/particle/trial1 ./basement/particle/trial2 ./basement/particle/trial3 ./basement/particle/trial4 ./basement/particle/trial5
+mkdir -p ./synthetic/particle/trial1 ./synthetic/particle/trial2 ./synthetic/particle/trial3 ./synthetic/particle/trial4 ./synthetic/particle/trial5
+
+# # populate the serialized datastructures for visualization
+../../build/bin/range_lib --method="cddt" --map_path=./basement_hallways_5cm.png --cddt_save_path=./basement/cddt.json
+../../build/bin/range_lib --method="pcddt" --map_path=./basement_hallways_5cm.png --cddt_save_path=./basement/pcddt.json
+../../build/bin/range_lib --method="cddt" --map_path=./synthetic.png --cddt_save_path=./synthetic/cddt.json
+../../build/bin/range_lib --method="pcddt" --map_path=./synthetic.png --cddt_save_path=./synthetic/pcddt.json
+
+# # grid benchmark for both maps/all methods
+../../build/bin/range_lib --method="rm,cddt,pcddt,bl,glt" --map_path=./basement_hallways_5cm.png --log_path=./basement/grid/trial1/ --which_benchmark=grid
+../../build/bin/range_lib --method="rm,cddt,pcddt,bl,glt" --map_path=./basement_hallways_5cm.png --log_path=./basement/grid/trial2/ --which_benchmark=grid
+../../build/bin/range_lib --method="rm,cddt,pcddt,bl,glt" --map_path=./basement_hallways_5cm.png --log_path=./basement/grid/trial3/ --which_benchmark=grid
+
[…]

### 2017-02-25 paper/data/make_statistics.py

@@ -1,0 +1,148 @@
+
+''' Makes a table like this:
+Method Name   
+-----------------------------------------------------------
+CDDT
+PCDDT
+LUT
+Ray Marching
+Bresenham's
+'''
+
+from tabulate import tabulate
+import csv
+import scipy
+import scipy.stats
+import numpy as np
+from scipy.ndimage.morphology import distance_transform_edt
+import matplotlib.pyplot as plt
+from matplotlib import cm
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+from numpy import genfromtxt
+
+BYTES_TO_MB = 1.0 / (1024.0*1024.0)
+
+class Trial(object):
+	""" Loads data from a single trial """
+	def __init__(self, path,method):
+		full_path = path + "/" + method + ".csv"
+		# print "Loading", full_path
+		self.methods = {}
+		self.raw_data = csv.reader(open(full_path,"rb"),delimiter=',')
+
+		self.data_header = next(self.raw_data)
+
+		if len(self.data_header) == 1:
+			self.data = np.array(list(self.raw_data)).astype('float')
+			# anything longer than this is an outlier, possibly because of a kernel interupt
+			self.data = self.data[self.data<0.0000040]
+			# self.data = self.data[self.data<0.0001]
+		else:
+			self.data = np.array(list(self.raw_data)).astype('float')[:,3]#*1000000000.0
+			# anything longer than this is an outlier, possibly because of a kernel interupt
+			self.data = self.data[self.data<0.0000040]
+			# self.data = self.data[self.data<0.0001]
+
+		self.construction_time = None
+		self.memory = None
+		try:
+			self.summary   = list(csv.reader(open(path + "/summary.csv","rb"),delimiter=','))
+			self.summary = filter(lambda x: method == x[0], self.summary)[0]
+
+			self.construction_time = float(self.summary[1])
+			self.memory = float(self.summary[2])*BYTES_TO_MB
+		except:
+			pass
+	def iqr(self):
+		q75, q25 = np.percentile(self.data, [75 ,25])
+		return q75 - q25
+
+	def mean(self):
+		return np.mean(self.data)
+
+	def median(self):
+		return np.median(self.data)
+
+	def hist(self):
+		plt.hist(self.data,bins=200)
+
+	def init_time(self):
+		return self.construction_time
+
+	def memory(self):
[…]

### 2017-02-25 paper/data/make_violin.py

@@ -1,0 +1,91 @@
+import csv
+import numpy as np
+from scipy.ndimage.morphology import distance_transform_edt
+import matplotlib.pyplot as plt
+from matplotlib import cm
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
+def reject_outliers(data, m = 12.):
+    d = np.abs(data - np.median(data))
+    mdev = np.median(d)
+    s = d/mdev if mdev else 0.
+    return data[s<m]
+
+def get_data(fn):
+	# print fn
+	reader = csv.reader(open(fn,"rb"),delimiter=',')
+
+	data_header = next(reader)
+
+	if len(data_header) == 1:
+		data = np.array(list(reader)).astype('float')
+		# anything longer than this is an outlier, possibly because of a kernel interupt
+		data = data[data<0.0000040]
+	else:
+		data = np.array(list(reader)).astype('float')[:,3]#*1000000000.0
+		# anything longer than this is an outlier, possibly because of a kernel interupt
+		data = data[data<0.0000040]
+
+
+	# raw_data = list(reader)
+	# # print raw_data[0], ra
+	# result=np.array(raw_data[1:]).astype('float')
+	# data = result[:,3]
+	# # data = reject_outliers(data)
+	# data = data[data<0.0000035]
+	data *= 1000 # convert to millis for better rendering
+	return data
+
+def violin_plot(path):
+	files = ["glt.csv", "pcddt.csv", "cddt.csv", "rm.csv", "bl.csv"]
+	label = ['LUT','PCDDT','CDDT','RM',"BL"]
+	files = map(lambda x: path + x, files)
+	# print files
+	data = map(get_data, files)
+
+	pos = [1,2,3,4,5]
+
+	# for i in xrange(5):
+	# 	print files[i]
+	# 	print np.mean(data[i])
+	# 	print np.median(data[i])
+	# 	print np.std(data[i])
+	# 	pass
+
+	# data = get_data("../range_libc/info/logs/1_lut.csv")
+	# pos = [1]
+
+	# reader = csv.reader(open("../range_libc/somefile.csv","rb"),delimiter=',')
+	# raw_data = list(reader)
+	# print raw_data[0]
+	# result=np.array(raw_data[1:]).astype('float')
+	# data = reject_outliers(result[:,3])
+	# print np.mean(result[:,3])
+
+	fig, axes = plt.subplots(1,1)
+
+	# # pos = [1, 2, 4, 5, 7, 8]
[…]

### 2017-02-25 visualize.py

@@ -158,6 +158,9 @@ class SliceScroller(object):
         plt.tight_layout()
         self.ax1.cla()
         self.ax2.cla()
+
+        self.ax1.axis('off')
+
         if not isinstance(self.ddts[self.ind], np.ndarray):
         # if self.ddts[self.ind] == None:
             self.ddts[self.ind] = np.sqrt(self.cddt.slices[self.ind].make_ddt(True))
@@ -165,12 +168,12 @@ class SliceScroller(object):
         ys = map(len, self.cddt.slices[self.ind].zeros)
         compression_factor = 2*self.cddt.map.width * self.cddt.map.height / (sum(ys))
 
-        self.ax1.set_title("DDT - Reconstructed from a slice of the CDDT, compression factor: " + str(compression_factor))
+        self.ax1.set_title("DDT - Reconstructed from a slice of the PCDDT, compression factor: " + str(compression_factor))
         self.ax1.set_ylabel('Theta = %s' % self.cddt.slices[self.ind].theta)
         self.ax1.imshow(self.ddts[self.ind],cmap="gray",interpolation='nearest', aspect='auto')
         
        
-        self.ax2.set_title("Number of entries in each CDDT bin")
+        self.ax2.set_title("Number of entries projected into each PCDDT bin")
         self.ax2.plot(ys)
         self.fig.canvas.draw()
         # self.im.set_data(self.ddt)

### 2017-02-25 paper/main.tex
updated conclusions to match measured data

@@
 This work demonstrates that the proposed CDDT algorithm may be used in mobile robotics to accelerate sensor model computation when localizing in a two dimensional occupancy grid map.
-While the precomputed LUT approach is generally 1.5 to 4 times faster than the proposed algorithm, the memory footprint of the proposed data structure is significantly smaller which may be desirable for resource constrained systems.
-Other than the LUT approach, CDDT is significantly faster than the other methods evaluated, typically ranging from a factor of 2 to 20.
+While the precomputed LUT approach is generally 1.1 to 1.7 times faster than the proposed algorithm, the memory footprint of the proposed data structure is significantly smaller which may be desirable for resource constrained systems.
+Other than the LUT approach, CDDT is significantly faster than the other methods evaluated.
+Ray marching is the next best, ranging from 1.28 to 2.25 times slower than PCDDT depending on access patterns and environmental factors.
+The comparison with the widely used Bresenham's Line algorithm is more stark, ranging from a factor of 5.39 to 14.8 speedup in our benchmarks.
 

### 2017-02-25 paper/main.tex
layed out figures nicely

@@
 \usepackage{tabu}
+\usepackage{float}
 % \usepackage{graphics} % for pdf, bitmapped graphics files
@@
 
-\begin{figure}[h]
-\img{synthetic_cropped.png}
-\caption{Occupancy grid map for the synthetic map example.}
-\label{violin:basement:grid}
-\end{figure}
-
 We have implemented the proposed algorithm in C++, as well as Bresenham’s Line, ray marching, and the LUT approach for comparison.
@@
 
-\subsection{Synthetic Benchmarks}
+% \begin{figure}[h]
+% \img{grid_trial2_basement_cropped.png}
+% \caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method.
+Basement map.
+X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
+% \label{violin:basement:grid}
+% \end{figure}
 
-\begin{figure}[h]
-\img{grid_trial2_basement_cropped.png}
+\begin{figure}[ht]
+\img{basement_grid_random_cropped.png}
+% \begin{center}\includegraphics[width=248px,height=520px]{basement_benchmarks_cropped.png}\end{center}
 \caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method.
@@
 X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
-\label{violin:basement:grid}
+\label{violin:basement:all}
 \end{figure}
 
-\begin{figure}[h]
-\img{grid_trial2_synthetic_cropped.png}
-\caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method.
+\subsection{Synthetic Benchmarks}
+
+% \begin{figure}[h]
+% \img{grid_trial2_basement_cropped.png}
+% \caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method.
+Basement map.
+X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
+% \label{violin:basement:grid}
+% \end{figure}
+
+% \begin{figure}[h]
+% \img{grid_trial2_synthetic_cropped.png}
+% \caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method.
 Synthetic map.
 X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
-\label{violin:basement:grid}
-\end{figure}
+% \label{violin:basement:grid}
+% \end{figure}
 
-\begin{figure}[h]
-\img{random_trial2_basement_cropped.png}
-\caption{Violin plots for a large number of randomly sampled queries. Basement map. The random access pattern degrades the caching performance variably for each method.}
-\label{violin:basement:random}
-\end{figure}
+% \begin{figure}[h]
+% \img{random_trial2_basement_cropped.png}
+% \caption{Violin plots for a large number of randomly sampled queries. Basement map. The random access pattern degrades the caching performance variably for each method.}
+% \label{violin:basement:random}
+% \end{figure}
 
-\begin{figure}[h]
-\img{random_trial2_synthetic_cropped.png}
-\caption{Violin plots for a large number of randomly sampled queries. Synthetic map. The random access pattern degrades the caching performance variably for each method.}
-\label{violin:basement:random}
-\end{figure}
+% \begin{figure}[h]
+% \img{random_trial2_synthetic_cropped.png}
+% \caption{Violin plots for a large number of randomly sampled queries. Synthetic map. The random access pattern degrades the caching performance variably for each method.}
+% \label{violin:basement:random}
+% \end{figure}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \begin{center}
@@
 \end{center}
-\caption{The construction time and memory footprint of each method for the Basement map.}
-\label{table:basement:init}
+\caption{The construction time and memory footprint of each method for the Synthetic map.}
+\label{table:synthetic:init}
 \end{figure}
@@
 \subsection{Particle Filter Experiments}
+
+\begin{figure}[h]
+\img{particle_trial2_basement_cropped.png}
+\caption{Violin plots for ray casting completion times during the execution of a particle filter.
+Basement map.
+Ray casting is primarily done inside of hallways, reducing the mean ray length and thereby improving Bresenham's Line performance.
+CDDT and PCDDT exploiting radial symmetry while resolving the sensor model.}
+\label{violin:basement:random}
+\end{figure}
 
@@
 
-\begin{figure}[h]
-\img{particle_trial2_basement_cropped.png}
-\caption{Violin plots for ray casting completion times during the execution of a particle filter.
-Basement map.
-Ray casting is primarily done inside of hallways, reducing the mean ray length and thereby improving Bresenham's Line performance.
-CDDT and PCDDT exploiting radial symmetry while resolving the sensor model.}
-\label{violin:basement:random}
-\end{figure}
-
 \section{CONCLUSIONS}
@@
 Ray marching is the next best, ranging from 1.28 to 2.25 times slower than PCDDT depending on access patterns and environmental factors.
-The comparison with the widely used Bresenham's Line algorithm is more stark, ranging from a factor of 5.39 to 14.8 speedup in our benchmarks.
+The comparison with the widely used Bresenham's Line algorithm is more stark, ranging from a factor of 5.39 to 14.8 in our benchmarks.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
+\img{synthetic_cropped.png}
+\caption{Synthetic map occupancy grid. 1024x1024}
+\label{violin:basement:grid}
+\end{figure}
+
+\begin{figure}[h!]
 \begin{center}
-\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
+\newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}}
+\begin{tabular}{ | C{1.15cm} | C{1.25cm}| C{1.25cm} | C{1.25cm} | C{1.25cm}|  } 
 \hline
-\multicolumn{5}{|c|}{Basement Map, Random Sampling Ray Cast} \\
+\multicolumn{5}{|c|}{\textbf{Synthetic Map Ray Cast Benchmarks}} \\
+\hline \hline
+\multicolumn{5}{|c|}{Random Sampling} \\
 \hline
 Method & Mean & Median & IQR & Speedup \\
+\hline
+BL & 1.19e-06 & 1.41e-06 & 7.71e-07 & 1 \\
+RM & 1.52e-07 & 1.25e-07 & 1.05e-07 & 7.81 \\
+CDDT & 1.24e-07 & 1.05e-07 & 5.3e-08 & 9.59 \\
+PCDDT & 1.19e-07 & 1.01e-07 & 5e-08 & 10.02 \\ 
+LUT & 1.82e-07 & 1.68e-07 & 1.4e-08 & 6.55 \\
+\hline \hline
+\multicolumn{5}{|c|}{Grid Sampling} \\
+\hline
+Method & Mean & Median & IQR & Speedup \\
+\hline
+BL & 1.03e-06 & 1.20e-06 & 6.79e-07 & 1 \\ 
+RM & 1.27e-07 & 1.03e-07 & 1.06e-07 & 8.06 \\
+CDDT & 7.02e-08 & 6.8e-08 & 1e-08 & 14.63 \\
+PCDDT & 6.94e-08 & 6.8e-08 & 9e-09 & 14.80 \\
+LUT & 6.33e-08 & 4.2e-08 & 4.6e-08 & 16.21 \\
+\hline
+\end{tabular}
+\caption{All times listed in seconds, speedup relative to Bresenham's Line}
+\label{table:basement:init}
+\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+\newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}}
+\begin{tabular}{ | C{1.15cm} | C{1.25cm}| C{1.25cm} | C{1.25cm} | C{1.25cm}|  } 
+\hline
+\multicolumn{5}{|c|}{\textbf{Basement Map Ray Cast Benchmarks}} \\
+\hline \hline
+Method & Mean & Median & IQR & Speedup \\
+\hline
+\multicolumn{5}{|c|}{Random Sampling} \\
 \hline
@@
 LUT & 1.89e-07 & 1.7e-07 & 2.1e-08 & 5.10 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{The construction time and memory footprint of each method for the Basement map.}
-\label{table:basement:init}
-\end{figure}
-
-\begin{figure}[h]
-\begin{center}
-\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
-\hline
-\multicolumn{5}{|c|}{Basement Map, Grid Sampling Ray Cast} \\
+\hline \hline
+\multicolumn{5}{|c|}{Grid Sampling} \\
 \hline
@@
 LUT & 6.13e-08 & 4.3e-08 & 4.6e-08 & 13.53 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{The construction time and memory footprint of each method for the Basement map.}
-\label{table:basement:init}
-\end{figure}
-
-\begin{figure}[h]
-\begin{center}
-\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
-\hline
-\multicolumn{5}{|c|}{Basement Map, Particle Filter Ray Cast} \\
+\hline \hline
+\multicolumn{5}{|c|}{Particle Filter} \\
 \hline
@@
 \end{tabular}
+
+\caption{All times listed in seconds, speedup relative to Bresenham's Line}
+\label{table:basement:init}
 \end{center}
-\caption{The construction time and memory footprint of each method for the Basement map.}
-\label{table:basement:init}
 \end{figure}
 
-\begin{figure}[h]
-\begin{center}
-\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
-\hline
-\multicolumn{5}{|c|}{Synthetic Map, Random Sampling Ray Cast} \\
-\hline
-Method & Mean & Median & IQR & Speedup \\
-\hline
-BL & 1.19e-06 & 1.41e-06 & 7.71e-07 & 1 \\
-RM & 1.52e-07 & 1.25e-07 & 1.05e-07 & 7.81 \\
-CDDT & 1.24e-07 & 1.05e-07 & 5.3e-08 & 9.59 \\
-PCDDT & 1.19e-07 & 1.01e-07 & 5e-08 & 10.02 \\ 
-LUT & 1.82e-07 & 1.68e-07 & 1.4e-08 & 6.55 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{The construction time and memory footprint of each method for the Basement map.}
-\label{table:basement:init}
-\end{figure}
-
-\begin{figure}[h]
-\begin{center}
-\begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
-\hline
-\multicolumn{5}{|c|}{Synthetic Map, Grid Sampling Ray Cast} \\
-\hline
-Method & Mean & Median & IQR & Speedup \\
-\hline
-BL & 1.03e-06 & 1.20e-06 & 6.79e-07 & 1 \\ 
-RM & 1.27e-07 & 1.03e-07 & 1.06e-07 & 8.06 \\
-CDDT & 7.02e-08 & 6.8e-08 & 1e-08 & 14.63 \\
-PCDDT & 6.94e-08 & 6.8e-08 & 9e-09 & 14.80 \\
-LUT & 6.33e-08 & 4.2e-08 & 4.6e-08 & 16.21 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{The construction time and memory footprint of each method for the Basement map.}
-\label{table:basement:init}
-\end{figure}
+% \begin{figure}[h!]
+% \begin{center}
+% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
+% \hline
+% \multicolumn{5}{|c|}{Basement Map, Random Sampling Ray Cast} \\
+% \hline
+% Method & Mean & Median & IQR & Speedup \\
+% \hline
+% BL & 9.66e-07 & 1.05e-06 & 1.08e-06 & 1 \\
+% RM & 2.12e-07 & 1.68e-07 & 1.64e-07 & 4.56 \\
+% CDDT & 1.58e-07 & 1.49e-07 & 9.1e-08 & 6.13 \\
+% PCDDT & 1.41e-07 & 1.32e-07 & 6.5e-08 & 6.83 \\
+% LUT & 1.89e-07 & 1.7e-07 & 2.1e-08 & 5.10 \\
+% \hline
+% \end{tabular}
+% \end{center}
+% \caption{Statistics per ray cast query for random sampling synthetic map.}
+% \label{table:basement:init}
+% \end{figure}
+% \begin{figure}[h!]
+% \begin{center}
+% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
+% \hline
+% \multicolumn{5}{|c|}{Basement Map, Grid Sampling Ray Cast} \\
+% \hline
+% Method & Mean & Median & IQR & Speedup \\
+% \hline
+% BL & 8.29e-07 & 9.24e-07 & 9.53e-07 & 1 \\
+% RM & 1.65e-07 & 1.26e-07 & 1.34e-07 & 5.02 \\
+% CDDT & 7.69e-08 & 7.2e-08 & 1.6e-08 & 10.78 \\
+% PCDDT & 7.32e-08 & 7e-08 & 1.4e-08 & 11.33 \\
+% LUT & 6.13e-08 & 4.3e-08 & 4.6e-08 & 13.53 \\
+% \hline
+% \end{tabular}
+% \end{center}
+% \caption{Statistics per ray cast query for grid sampling synthetic map.}
+% \label{table:basement:init}
+% \end{figure}
+% \begin{figure}[h!]
+% \begin{center}
+% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.15cm}|  } 
+% \hline
+% \multicolumn{5}{|c|}{Basement Map, Particle Filter Ray Cast} \\
+% \hline
+% Method & Mean & Median & IQR & Speedup \\
+% \hline
+% BL & 3.89e-07 & 2.67e-07 & 2.31e-07 & 1 \\
+% RM & 1.56e-07 & 1.18e-07 & 9.4e-08 & 2.49 \\
+% CDDT & 7.67e-08 & 7.8e-08 & 5.7e-08 & 5.07 \\
+% PCDDT & 7.22e-08 & 7.4e-08 & 5.4e-08 & 5.39 \\
+% LUT & 4.35e-08 & 4.1e-08 & 5e-09 & 8.94 \\
+% \hline
+% \end{tabular}
+% \end{center}
+% \caption{Statistics per ray cast query for random sampling basement map.}
+% \label{table:basement:init}
+% \end{figure}
+% \begin{figure}[h!]
+% \begin{center}
+% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.15cm}|  } 
+% \hline
+% \multicolumn{5}{|c|}{Synthetic Map, Random Sampling Ray Cast} \\
+% \hline
+% Method & Mean & Median & IQR & Speedup \\
+% \hline
+% BL & 1.19e-06 & 1.41e-06 & 7.71e-07 & 1 \\
+% RM & 1.52e-07 & 1.25e-07 & 1.05e-07 & 7.81 \\
+% CDDT & 1.24e-07 & 1.05e-07 & 5.3e-08 & 9.59 \\
+% PCDDT & 1.19e-07 & 1.01e-07 & 5e-08 & 10.02 \\ 
+% LUT & 1.82e-07 & 1.68e-07 & 1.4e-08 & 6.55 \\
+% \hline
+% \end{tabular}
+% \end{center}
+% \caption{The construction time and memory footprint of each method for the Basement map.}
+% \label{table:basement:init}
+% \end{figure}
+% \begin{figure}[h!]
+% \begin{center}
+% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.15cm}|  } 
+% \hline
+% \multicolumn{5}{|c|}{Synthetic Map, Grid Sampling Ray Cast} \\
+% \hline
+% Method & Mean & Median & IQR & Speedup \\
+% \hline
+% BL & 1.03e-06 & 1.20e-06 & 6.79e-07 & 1 \\ 
+% RM & 1.27e-07 & 1.03e-07 & 1.06e-07 & 8.06 \\
+% CDDT & 7.02e-08 & 6.8e-08 & 1e-08 & 14.63 \\
+% PCDDT & 6.94e-08 & 6.8e-08 & 9e-09 & 14.80 \\
+% LUT & 6.33e-08 & 4.2e-08 & 4.6e-08 & 16.21 \\
+% \hline
+% \end{tabular}
+% \end{center}
+% \caption{The construction time and memory footprint of each method for the Basement map.}
+% \label{table:basement:init}
+% \end{figure}
 
@@
 \end{thebibliography}
+
 \end{document}

### 2017-02-27 paper/main.tex
several comments

@@
 
+%: Margin note commands
+ \usepackage[usenames,dvipsnames]{xcolor}
+\setlength{\marginparwidth}{0.5in}
+\let\oldmarginpar\marginpar
+\renewcommand\marginpar[1]{\-\oldmarginpar[\raggedleft\footnotesize #1]%
+{\raggedright\footnotesize #1}}
+\newcommand{\skcomment}[2]{{\color{red}#1} \marginpar{\color{TealBlue}\tiny #2 (SK)}}
+
 % \begin{center}\frame{\includegraphics[width=\columnwidth]{synthetic.png}}\end{center}
@@
 % \thanks{*This work was not supported by any organization}% <-this % stops a space
-\thanks{$^{1}$Corey H. Walsh is with the Department of Computer Science and Engineering, 
-Albert Author is with Faculty of Electrical Engineering, Mathematics and Computer Science, Massachusetts Institute of Technology,
+\thanks{$^{1}$Corey H. Walsh is with the Department of Computer Science and Engineering, Massachusetts Institute of Technology,
         Cambridge, MA 02139, USA
         {\tt\small chwalsh@mit.edu}}%
-\thanks{$^{2}$Sertac Karaman is with Faculty of Computer Science and Engineering, Massachusetts Institute of Technology,
+\thanks{$^{2}$Sertac Karaman is with the Laboratory for Information and Decision Systems, Massachusetts Institute of Technology,
         Cambridge, MA 02139, USA
@@
 \pagestyle{empty}
+
+\skcomment{}{I would consider putting the word ``localization'' in the title.}
 
@@
 
+\skcomment{}{Perhaps start with a short introduction to localization, and then describe what particle filter (Monte Carlo) localization is, and then introduce ray casting in that context.}
+
 The ray cast operation is an important sub-problem for robot localization \cite{localization} using distance sensors to detect the environment.
@@
 
-To combat this challenge, Fox et al.
-\cite{localization} suggest the use of a large three-dimensional lookup table (LUT) to store expected ranges for each discrete $(x,y,\theta)$ state.
-While this is simple to implement and does result in large speed improvements as compared to ray casting, it can be prohibitively memory intensive for large maps and/or resource constrained systems.
+To combat this challenge, \skcomment{Fox et al.}{Thrun et al.?} \cite{localization} suggest the use of a large three-dimensional lookup table (LUT) to store expected ranges for each discrete $(x,y,\theta)$ state.
+While this is simple to implement and does result in large speed improvements as compared to ray casting, it can be prohibitively memory intensive for large maps and/or resource constrained systems\skcomment{.}{perhaps give an example: In such and such large area with such and such resolution map, the data structure is XX GigaBytes (TeraBytes?)}
+
 We propose an analogous data structure called a Compressed Directional Distance Transform (CDDT) which requires significantly less memory and precomputation time while still allowing near constant time ray casting queries.
 
-Many robotic applications require three dimensional maps, however that is considered outside the scope of this work since many data structures exist for accelerating 3D ray casting, such as the bounded volume hierarchy and related variants.
-The five dimensional table required to precompute expected ranges for a discrete state space $(x,y,z,\theta,\phi)$ would be prohibitively expensive to both compute and store, so 3D acceleration data structures are exclusively used when 3D ray casting is required.
-Though our approach does extend to higher dimensionality and would be smaller than the analogous five dimensional table, it would still be prohibitively expensive in greater than two dimensions, thus we do not consider it here.
+\skcomment{}{This is probably a good place to list the contributions of this work as in: ``The contributions of this work are the following.
+Firstly, we develop ...
+Secondly, we implement this algorithm for a ...
+You can also say things like: We observe two orders of magnitude improvement in memory requirements, with little sacrifice in computation time, when compared to lookup table methods.
+Et cetera.}
 
+\skcomment{Many robotic applications require three dimensional maps, however that is considered outside the scope of this work since many data structures exist for accelerating 3D ray casting, such as the bounded volume hierarchy and related variants.}{I would put the discussion in this paragraph to the future work section in the conclusions, rather than here.} The five dimensional table required to precompute expected ranges for a discrete state space $(x,y,z,\theta,\phi)$ would be prohibitively expensive to both compute and store, so 3D acceleration data structures are exclusively used when 3D ray casting is required.
+\skcomment{Though our approach does extend to higher dimensionality and would be smaller than the analogous five dimensional table, it would still be prohibitively expensive in greater than two dimensions, thus we do not consider it here.}{I would not worry about putting this justification either.}
+
+\skcomment{
 \section{RELATED WORK}
-
+}{You might want to integrate this section into the introduction section.
+Otherwise the introduciton section is unusually short.
+Many people discuss related work in the introduciton.
+You can put this discussion right after your second paragraph, and right before listing all your contributions.}
 Several widely used algorithms exist for ray casting in a two dimensional space.
@@
 
-\subsection{Bresenham's Line}
+\skcomment{\subsection{Bresenham's Line}}{I would not worry about these section headings. You say the names of the algorithms in the beginning of the paragraphs anyways.}
 
@@
 For example, a query ray which travels closely parallel to a wall will make small steps between each query point for the whole length of the ray.
-As a result, the performance of ray marching has a long tail distribution which can be problematic for near real time algorithms.
+As a result, the performance of ray marching \skcomment{has a long tail distribution}{is this widely known?
+Is there a paper that you can cite here?
+Or can you refer the reader to your results in this paper (e.g., write ``(See Section X.)'' to tell the reader that there are some results there on heavy tails of this algorithm.} which can be problematic for near real time algorithms.
 Though highly scene dependent, on average ray marching runtime is roughly logarithmic in distance between obstacles, though in the worst case it is linear in map size.
@@
 
+\skcomment{
 \section{APPROACH}
+}{Instead of this section title, i would use ``The Compressed Directional Distance Transform (CDDT) Algorithm'' or something like that.}
 
@@
 In contrast, to compute the DDT, the obstacles in the map are rotated about the center of the map by $-\theta_d$ and the values of each column in the DDT may be directly computed as a sawtooth function which goes to zero whenever an obstacle is encountered in that column.
-Thus, each column in the DDT may be characterized as a sawtooth function with uneven periodicity.
+\skcomment{}{Do you think this is a good place for an example?} Thus, each column in the DDT may be characterized as a sawtooth function with \skcomment{uneven periodicity}{what is uneven periodicity mean?}.
 While it is true that the rotation of the occupancy grid introduces small errors as compared to ray casting for each cell, as previously discussed small errors do not significantly impact particle filter localization performance due to the use of a probabilistic error model.
@@
 
-The conversion from a DDT to a CDDT is performed by converting the sawtooth response of each column into a set of points in which the sawtooth wave goes to zero.
-This conversion is lossless in the sense that each element of the table may be recovered as a function of the nearest zero point in a particular direction.
+\skcomment{The conversion from a DDT to a CDDT is performed by converting the sawtooth response of each column into a set of points in which the sawtooth wave goes to zero.}{This is probably the most important part, but kind of hard to understand.
+I would consider an example here as well.} This conversion is lossless in the sense that each element of the table may be recovered as a function of the nearest zero point in a particular direction.
 If the set of zero points is sorted, then the magnitude of the saw tooth wave at any point in the original DDT may be quickly discovered by performing binary search for the nearest zero point in that column and finding the distance between the query point and that zero point.
@@
 Once all scene geometry is projected into the table, each list of zero points is sorted to facilitate indexing.
+
+\skcomment{}{This section could be organized a bit better.
+Perhaps have a subsection that says ``Compression algorithm" (text above here) and another one that says ``Query Algorithm'' (text below here).}
 
@@
 To ease this burden, we provide our implementation as well as Python wrappers with an Apache 2.0 license.
+
+\skcomment{}{You can put all the text after this point into one subsection, which you can call ``Further optimizations'' or something like that.}
 
@@
 
-In this case, the cost of inserting or deleting an occupied grid cell would be $O(theta\_discretization*\log(number of lut bin elements))$.
-The number of elements in each lut bin tends to be small and is bounded, so this becomes $O(theta\_discretization)$ which is likely not prohibitive for real-time performance.
+In this case, the cost of inserting or deleting an occupied grid cell would be \skcomment{$O(theta\_discretization*\log(number of lut bin elements))$.}{I think this is your plan already, but just to note: Please make sure to use symbols for these and introduce what the symbols mean.
+} The number of elements in each lut bin tends to be small and is bounded, so this becomes $O(theta\_discretization)$ which is likely not prohibitive for real-time performance.
 
@@
 By removing entries in the CDDT which can never possibly result in a ray collision, it is possible to further reduce the memory footprint of the data structure.
-Consider a 3x3 block of obstacles in an otherwise empty map.
-The center obstacle will never be the nearest neighbor in any ray casting query, because any such query would first intersect with one of the 8 surrounding obstacles.
+\skcomment{Consider a 3x3 block of obstacles in an otherwise empty map.}{I think this deserves a figure, just to allow the reader to quickly read your paper.} The center obstacle will never be the nearest neighbor in any ray casting query, because any such query would first intersect with one of the 8 surrounding obstacles.
 To exploit this, one can take the morphological edge map of the occupancy grid map prior to CDDT generation without loss of generality.
@@
 
-Additionally, consider a line of obstacles aligned along the Y-axis.
-Every element in this line will be projected into a single list of zero points for the theta=0 slice of the CDDT.
+\skcomment{Additionally, consider a line of obstacles aligned along the Y-axis.}{While you are at it, put a figure for this one too?
+} Every element in this line will be projected into a single list of zero points for the theta=0 slice of the CDDT.
 However, the middle elements of the line will never result in a collision.
@@
 
+\skcomment{}{Do you think you can have a subsection called ``analysis'' which has a few propositions/theorems that say things like Memory requirement is $O(NM)$ and running time is $O(1)$?
+This may be hard given the deadline, but osmething to think about.
+}
+
 \section{EXPERIMENTS}
 
-We have implemented the proposed algorithm in C++, as well as Bresenham’s Line, ray marching, and the LUT approach for comparison.
+We have implemented the proposed algorithm in the C++ programming language, as well as Bresenham’s Line, ray marching, and the LUT approach for comparison.
 Our source code is available for use and analysis [TODO add a link to github], and Python wrappers are also provided for easier usage.

### 2017-02-27 paper/main.tex
made large progress on initial editing pass

@@
 %\usepackage{times} % assumes new font selection scheme installed
-%\usepackage{amsmath} % assumes amsmath package installed
+\usepackage{amsmath} % assumes amsmath package installed
 %\usepackage{amssymb}  % assumes amsmath package installed
+\usepackage{tikz}
+\usetikzlibrary{matrix}
+\usetikzlibrary{shapes.geometric, arrows}
 
@@
 \newcommand{\imgFrame}[1]{\begin{center}\frame{\includegraphics[width=\columnwidth]{{#1}}}\end{center}}
+
+\newcommand{\gridScale}{1.25}
 
@@
 {\raggedright\footnotesize #1}}
-\newcommand{\skcomment}[2]{{\color{red}#1} \marginpar{\color{TealBlue}\tiny #2 (SK)}}
+\newcommand{\skcomment}[2]{{\color{red}#1} \marginpar{\color{blue}\tiny #2 (SK)}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{INTRODUCTION}
+\section{Introduction}
 
-\skcomment{}{Perhaps start with a short introduction to localization, and then describe what particle filter (Monte Carlo) localization is, and then introduce ray casting in that context.}
+Determining a robot's location and orientation in a known environment, also known as localization, is an important and challenging problem in the field of robotics.
+Particle filters are a popular class of Monte Carlo algorithms used to track the pose of mobile robots by iteratively refining a set of pose hypotheses called particles.
+After determining an initial set of particles, the particle filter updates the position and orientation of each particle by applying a movement model based on available odometry data.
+Next, the belief in each particle is updated by comparing sensor readings to a map of the environment.
+Finally, the particles are resampled according to the belief distribution and algorithm repeats.
 
-The ray cast operation is an important sub-problem for robot localization \cite{localization} using distance sensors to detect the environment.
-In particle filter localization, particles are re-weighted according to a Bayesian update equation which relies on ray casting to determine the ground truth distance between the hypothesis pose and obstacles in a map.
-Effective particle filters often maintain thousands of particles, each of which must be reweighted tens of times per second, and each re-weighting may require tens of ray casts.
-As a result, millions of ray cast operations may be resolved per second, posing a significant computational challenge for resources constrained systems.
+While particle filters provide a robust and general framework for solving the localization problem, they can be computationally expensive due to both the number of particles which must be maintained and the evaluation of the sensor model.
+In robots with range sensors such as LiDAR or Sonar, ray casting is frequently used to compare sensor readings with the ground truth distance between the hypothesis pose and obstacles in a map.
+Ray casting itself is a complex operation, and reweighting any single particle may require tens of ray casts.
+Many effective particle filters maintain thousands of particles, updating each particle tens of times per second.
+As a result, millions of ray cast operations may be resolved per second, posing a significant computational challenge for resource constrained systems.
 
-To combat this challenge, \skcomment{Fox et al.}{Thrun et al.?} \cite{localization} suggest the use of a large three-dimensional lookup table (LUT) to store expected ranges for each discrete $(x,y,\theta)$ state.
-While this is simple to implement and does result in large speed improvements as compared to ray casting, it can be prohibitively memory intensive for large maps and/or resource constrained systems\skcomment{.}{perhaps give an example: In such and such large area with such and such resolution map, the data structure is XX GigaBytes (TeraBytes?)}
+Several well known algorithms exist for ray casting in two dimensional spaces such as Bresenham's Line algorithm \cite{bresenham} and ray marching \cite{raymarching}.
+Both algorithms work by iteratively checking points in the map starting at the query point and moving in the ray direction until an obstacle is discovered.
+This process may require hundreds of memory reads per ray cast depending on the distance to the nearest obstacle, and does not provide constant time performance.
 
-We propose an analogous data structure called a Compressed Directional Distance Transform (CDDT) which requires significantly less memory and precomputation time while still allowing near constant time ray casting queries.
+To combat the computational challenges of ray casting while localizing in a two-dimensional map, Thrun et al.
+\cite{localization} suggest the use of a large three-dimensional lookup table (LUT) to store expected ranges for each discrete $(x,y,\theta)$ state.
+While this is simple to implement and does result in large speed improvements as compared to ray casting, it can be prohibitively memory intensive for large maps and/or resource constrained systems.
+In a 2000 by 2000 occupancy map, storing ranges for 200 discrete directions would require over 1.5GB.
+While this memory requirement may be acceptable in many cases, it scales with the square of map size - a 4000 by 4000 map would require over 6GB for the same angular discretization.
 
-\skcomment{}{This is probably a good place to list the contributions of this work as in: ``The contributions of this work are the following.
-Firstly, we develop ...
-Secondly, we implement this algorithm for a ...
-You can also say things like: We observe two orders of magnitude improvement in memory requirements, with little sacrifice in computation time, when compared to lookup table methods.
-Et cetera.}
+We propose an analogous data structure to the three-dimensional LUT called a Compressed Directional Distance Transform (CDDT) which requires significantly less memory and precomputation time while still allowing near constant time ray casting queries.
+The contributions of this work are the following.
+Firstly, we develop a problem formulation and provide details of our method.
+Secondly, we implement the proposed algorithm and compare its performance to alternate two-dimensional ray casting methods.
+Finally, we discuss potential extensions to the CDDT algorithm which could be the basis of future research.
+We observe two orders of magnitude improvement in memory requirements, with little sacrifice in computation time, when compared to the lookup table methods.
+Additionally, we observe a large speedup when compared to the other ray casting methods considered with similar memory requirements.
 
-\skcomment{Many robotic applications require three dimensional maps, however that is considered outside the scope of this work since many data structures exist for accelerating 3D ray casting, such as the bounded volume hierarchy and related variants.}{I would put the discussion in this paragraph to the future work section in the conclusions, rather than here.} The five dimensional table required to precompute expected ranges for a discrete state space $(x,y,z,\theta,\phi)$ would be prohibitively expensive to both compute and store, so 3D acceleration data structures are exclusively used when 3D ray casting is required.
-\skcomment{Though our approach does extend to higher dimensionality and would be smaller than the analogous five dimensional table, it would still be prohibitively expensive in greater than two dimensions, thus we do not consider it here.}{I would not worry about putting this justification either.}
+\section{Related Work}
 
-\skcomment{
-\section{RELATED WORK}
-}{You might want to integrate this section into the introduction section.
-Otherwise the introduciton section is unusually short.
-Many people discuss related work in the introduciton.
-You can put this discussion right after your second paragraph, and right before listing all your contributions.}
-Several widely used algorithms exist for ray casting in a two dimensional space.
 Due to varied handling of edge cases and ambiguities in the problem of determining distance between any point and the nearest obstacle in a particular direction, most 2D ray casting algorithms do not provide exactly consistent results for every query.
@@
 
-\skcomment{\subsection{Bresenham's Line}}{I would not worry about these section headings. You say the names of the algorithms in the beginning of the paragraphs anyways.}
-
-Bresenham's line algorithm \cite{bresenham} incrementally determines the set of pixels that approximate the trajectory of query ray starting from the query point $(x,y)$ and progressing in the $\theta$ direction.
+Bresenham's line algorithm \cite{bresenham} incrementally determines the set of pixels that approximate the trajectory of query ray starting from the query point $(x,y)$ and progressing in the $\theta$ direction one pixel at a time.
 The algorithm terminates once the nearest occupied pixel is discovered, and the euclidean distance between that occupied pixel and the query $(x,y)$ is reported.
@@
 
-\subsection{Ray Marching Distance Transforms}
-
 Similar to Bresenham's Line algorithm, ray marching \cite{raymarching} checks points along the line radiating outwards from the query point until an obstacle is encountered.
-The primary difference is that ray marching makes larger steps along the query ray, avoiding unnecessary memory reads.
+The primary difference is that ray marching makes larger steps along the query ray, thereby avoiding unnecessary memory reads.
 Beginning at the query point $(x,y)$ the ray marching algorithm proceeds in the theta direction, stepping along the line by the minimum distance between each query point and the nearest obstacle.
-The algorithm terminates when the query point falls on an obstacle in the map.
-A precomputed euclidean distance transform of the map provides the distance between each query point and the nearest obstacle in any direction.
+The algorithm terminates when the query point coincides with an obstacle in the map.
+A precomputed euclidean distance transform of the occupancy map provides the distance between each query point and the nearest obstacle in any direction.
 
@@
 
-Ray marching is on average much faster than Bresenham’s line, but edge cases exist in which the performance is equivalent.
-For example, a query ray which travels closely parallel to a wall will make small steps between each query point for the whole length of the ray.
-As a result, the performance of ray marching \skcomment{has a long tail distribution}{is this widely known?
-Is there a paper that you can cite here?
-Or can you refer the reader to your results in this paper (e.g., write ``(See Section X.)'' to tell the reader that there are some results there on heavy tails of this algorithm.} which can be problematic for near real time algorithms.
-Though highly scene dependent, on average ray marching runtime is roughly logarithmic in distance between obstacles, though in the worst case it is linear in map size.
+Ray marching is on average much faster than Bresenham’s line, but edge cases exist in which the performance is nearly equivalent.
+As noted in \cite{acceleratedraymarching} the traversal speed of rays rapidly decreases as sampled positions approach obstacles.
+For this reason, rays which travel parallel to walls progress very slowly as compared to those passing through open areas.
+Thus the performance of ray marching exhibits a long tail distribution as seen in (Fig.
+\ref{violin:basement:all}, \ref{violin:basement:particle}) which negatively impacts average runtime as compared to the median, and can be problematic for near real time algorithms.
 
-% \begin{figure}[thpb]
-%       \centering
-%       \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
-% }}
-%       %\includegraphics[scale=1.0]{figurefile}
-%       \caption{Inductance of oscillation winding on amorphous
-%       magnetic core versus DC bias magnetic field}
-%       \label{figurelabel}
-%   \end{figure}
-
-\subsection{Lookup Table}
-
-As previously described, a common acceleration technique for two dimensional ray casting is to simply precompute the ray distances for every state in a discrete grid and store the results in a three dimensional LUT.
+As previously described, a common acceleration technique for two dimensional ray casting is to precompute the ray distances for every state in a discrete grid and store the results in a three dimensional LUT.
 Any ray casting technique may be used to compute the table.
@@
 State space discretization implies approximate results, since intermediate states must be rounded onto the discrete grid.
-While the effect is small for x and y, it can be significant for theta.
+While the effect of rounding query position, it can be significant for $\theta$.
 As the ray moves further away from the query point along the theta direction, angular discretization error accumulates.
 For queries $(x,y,\theta)$ discretized into $(x_d,y_d,\theta_d)$, the distance between the end of the ray $(x_d,y_d,Θ_d)$ and its projection onto the line implied by $(x,y,\theta)$ becomes large as the length of the ray increases.
-Although the error induced by discretization may be unacceptable for computer graphics applications, it is generally acceptable for probabilistic algorithms such as the recursive Bayes filter since error is expected and directly modeled.
+Although the error induced by discretization may be unacceptable for computer graphics applications, it is generally acceptable for probabilistic algorithms such as the particle filter since error is expected and directly modeled.
 
-One method to improve accuracy of queries which fall between discrete states is to ray cast the neighboring discrete states and interpolate between them, rather than simply rounding to the nearest discrete state.
-Since we are evaluating the performance of these methods for use in a particle filter, we do not perform interpolation since the increased computation reduces the number of particles that may be maintained in real time and therefore any increase in accuracy with respect to the stored map is effectively negated.
+Rather than simply rounding to the nearest grid state, one method to improve accuracy of queries which lie between discrete states is to query the neighboring states and interpolate between them.
+Since we are evaluating the performance of these methods for use in a particle filter, we do not perform interpolation since the increased computation reduces the number of particles that may be maintained in real time, and therefore any increase in accuracy with respect to the stored map is effectively negated.
 
-\skcomment{
-\section{APPROACH}
-}{Instead of this section title, i would use ``The Compressed Directional Distance Transform (CDDT) Algorithm'' or something like that.}
+\section{Problem Formulation and Notation}
+
+We define the problem of ray casting in occupancy grids as follows.
+We assume a known occupancy grid map in which occupied cells have value one and unoccupied cells have value zero.
+Given a query pose $(x,y,\theta)_{query}$ in map space, the ray cast operation finds the nearest occupied pixel $(x,y)_{collide}$ which lies on the ray starting at the position $(x,y)_{query}$ pointing in the $\theta_{query}$ direction.
+The ray cast operation returns the euclidean distance $d_{ray}$ between $(x,y)_{query}$ and $(x,y)_{collide}$.
+
+$$ d_{ray} = \left\Vert\begin{pmatrix}
+           x \\
+           y 
+         \end{pmatrix}_{query} - 
+         \begin{pmatrix}
+           x \\
+           y
+         \end{pmatrix}_{collide}\right\Vert_2
+$$
+
+We refer to $(x,y,\theta)_{query}$ as the query pose, and $(x,y)_{query}$ as the query point.
+Additionally, denote the discretized query pose as $\lfloor(x,y,\theta)_{query}\rceil$.
+A $\theta$ slice through the LUT refers to all table entries for that particular $\lfloor\theta\rceil$.
+
+\begin{figure}[h]
+\img{coord_system.png}
+\caption{Occupancy grid map coordinate system}
+\label{coordinate_system}
+\end{figure}
+
+\section{The Compressed Directional Distance Transform Algorithm}
 
 Although the three dimensional table used to store precomputed ray cast solutions for a discrete state space is inherently large, it is highly compressible.
-This is most apparent in the cardinal directions, in which adjacent values along a particular dimension of the table increase by exactly one unit of distance for unobstructed positions.
+This is most apparent in the cardinal directions, in which adjacent values along a particular dimension of the table increase by exactly one unit of distance for unobstructed positions as in Fig.
+\ref{cardinal_example}.
 Our data structure is designed to compress this redundancy while still allowing for fast queries in near constant time.
-We accomplish this though through what we refer to as a compressed directional distance transform (CDDT) described here.
+We accomplish this though through what we refer to as a Compressed Directional Distance Transform (CDDT) described here.
 
 \begin{figure}[h]
-\img{basement_bw_cropped.png}
-\caption{Occupancy grid map of the Stata basement loop. 1200x1200}
-\label{stata}
+\begin{center}
+\begin{tikzpicture}
+\fill[black,scale=\gridScale] (0.5,0.5) rectangle (1,1);
+\fill[black,scale=\gridScale] (-0.5,0) rectangle (0,0.5);
+\fill[black,scale=\gridScale] (0,-0.5) rectangle (0.5,0);
+\fill[black,scale=\gridScale] (0.5,-1) rectangle (1,-0.5);
+\draw[scale=\gridScale,step=0.5cm,color=white] (-1,-1.1) grid (1,1.1);
+\draw[scale=\gridScale,step=0.5cm,color=gray] (-1,-1) grid (1,1);
+\end{tikzpicture}
+\begin{tikzpicture}
+\draw[scale=\gridScale,step=0.5cm,color=white] (-1,-1.1) grid (1,1.1);
+\draw[scale=\gridScale,step=0.5cm,color=gray] (-1,-1) grid (1,1);
+\matrix[matrix of nodes,nodes={inner sep=0pt,text width=.5cm,align=center,minimum height=0.45cm, scale=\gridScale}]{
+3 & 2 & 1 & 0 \\
+1 & 0 & - & - \\
+2 & 1 & 0 & - \\    
+3 & 2 & 1 & 0\\};
+\end{tikzpicture}
+\end{center}
+\caption{Occupancy grid (left) and associated LUT for $\theta=0$ (right).}
+\label{cardinal_example}
 \end{figure}
 
-The euclidean distance transform (EDT) of an occupancy grid map stores the distance to the nearest obstacle in any direction for each point $(x,y)$ in the map.
-In contrast, a directional distance transform (DDT) stores the distance to the nearest obstacle in a particular direction for each point $(x,y)$ in the map.
-The key difference between a single Θ slice of the LUT and a directional distance transform is the way in which it is computed.
-To compute the LUT slice, ray casting is performed in the $\theta_d$ direction from every $(x_d,y_d)$ in the map using another algorithm such as Bresenham’s line.
-In contrast, to compute the DDT, the obstacles in the map are rotated about the center of the map by $-\theta_d$ and the values of each column in the DDT may be directly computed as a sawtooth function which goes to zero whenever an obstacle is encountered in that column.
-\skcomment{}{Do you think this is a good place for an example?} Thus, each column in the DDT may be characterized as a sawtooth function with \skcomment{uneven periodicity}{what is uneven periodicity mean?}.
-While it is true that the rotation of the occupancy grid introduces small errors as compared to ray casting for each cell, as previously discussed small errors do not significantly impact particle filter localization performance due to the use of a probabilistic error model.
-In practice, errors induced by map rotation are dominated by those inherently present in distance sensors.
+The euclidean distance transform of an occupancy grid map stores the distance to the nearest obstacle in any direction $\theta$ for each possible $\lfloor(x,y)\rceil$ in the map.
+In contrast, a directional distance transform (DDT) stores the distance to the nearest obstacle in a particular direction $\lfloor\theta\rceil$ for every $\lfloor(x,y)\rceil$.
+The key difference between a single $\theta$ slice of the LUT and a directional distance transform for the same $\theta$ is the way in which it is computed and indexed.
+To compute the LUT slice, ray casting is performed in the $\theta$ direction for every $\lfloor(x,y)\rceil$.
+At runtime, each $(x,y)_{query}$ is discretized to $\lfloor(x,y)_{query}\rceil$ and the LUT slice is directly indexed.
 
-\begin{figure}[h]
-\img{slice0_cddt_basement.png}
-\caption{A single slice of the DDT generated from the CDDT data structure for $\theta=0$.
-The value of each pixel represents the distance to the nearest obstacle in the upwards direction with respect to the image above.
-Each column may be thought of as a sawtooth wave, and the points where the wave is zero are projected into the CDDT bin for that column for later indexing.
-The blue plot indicates the number of zero points projected into each CDDT bin.
-The compression factor of 267 is computed as two times the size of the corresponding LUT slice divided by the size of this slice in the CDDT.
-The factor of two is due to the use of rotational symmetry to reduce the number of slices maintained in the CDDT with respect to the LUT.}
-\label{cddt:slice0}
-\end{figure}
+In contrast, to compute the DDT, the obstacles in the map are rotated about the origin by $-\lfloor\theta\rceil$ and ray casting is implicitly performed in the $\theta=0$ direction to populate the table.
+Then, to index the table for a query $(x,y,\theta)_{query}$ one discretizes to $\lfloor(x,y,\theta)_{query}\rceil$, rotates $\lfloor(x,y)_{query}\rceil$ about the origin, and uses the rotated coordinates $\lfloor(x,y,\theta)_{query}\rceil_{rot}$ to index into the DDT.
+Thus, the same operation is computed in both the DDT and the LUT, but while one changes the ray cast direction to populate the LUT, one rotates scene geometry and ray casts in a constant direction to populate the DDT.
+While it is true that the rotation of the scene geometry introduces small errors as compared to ray casting for each cell, as previously discussed small errors do not significantly impact particle filter localization performance due to the use of a probabilistic error model and are dominated by errors inherently present in distance sensors.
 
-\begin{figure}[h]
-\img{slice41_cddt_basement.png}
-\caption{Another slice of the DDT generated from the CDDT data structure for $\theta=2.385$.}
-\label{cddt:slice41}
-\end{figure}
+The distinction between the LUT slice and the DDT may be subtle, but it has an important effect.
+Since ray casting is always performed in the $\theta=0$ direction to populate the DDT, all values in the same row of the DDT either increase by one unit with respect to their neighbor in the $\theta=0$ direction or go to zero.
+Thus each row of the DDT may be characterized as a sawtooth function where the zero points correspond to obstacles in the map.
+This characterization as a sawtooth function provides a natural method of lossless compression: keep the zero points and discard the rest.
 
-\skcomment{The conversion from a DDT to a CDDT is performed by converting the sawtooth response of each column into a set of points in which the sawtooth wave goes to zero.}{This is probably the most important part, but kind of hard to understand.
-I would consider an example here as well.} This conversion is lossless in the sense that each element of the table may be recovered as a function of the nearest zero point in a particular direction.
-If the set of zero points is sorted, then the magnitude of the saw tooth wave at any point in the original DDT may be quickly discovered by performing binary search for the nearest zero point in that column and finding the distance between the query point and that zero point.
-This losslessness and fast indexing allows for ray cast queries to be quickly resolved without storing the entire DDT, but rather by storing the sorted lists of zero points.
-For sake of efficiency, the full DDT need not be computed, but rather the CDDT may be built directly by projecting each obstacle into the coordinate space of each theta slice and storing a list of y-coordinates for each discrete x-coordinate.
-Once all scene geometry is projected into the table, each list of zero points is sorted to facilitate indexing.
+The conversion from a DDT to a CDDT slice is performed by storing the $x$ coordinate of each zero point in the coordinate space of the DDT for every discrete row.
+At query time, the sawtooth function (which is the distance to the nearest obstacle in the query direction) encoded in the DDT may be quickly recovered by performing binary search for nearest zero point in the correct row of the CDDT slice.
+We refer to the list of zero points for a single row as a CDDT bin.
+Similar to the LUT, the full CDDT is defined as every CDDT slice for all discrete values of $\theta$.
 
-\skcomment{}{This section could be organized a bit better.
-Perhaps have a subsection that says ``Compression algorithm" (text above here) and another one that says ``Query Algorithm'' (text below here).}
+For performance, it is not necessary to compute the full DDT, but rather the CDDT may be directly constructed by projecting each obstacle into the coordinate space of the DDT for every $\lfloor\theta\rceil$ and sorting its $x$ coordinate in the CDDT bin corresponding to its $y$ coordinate and rotation.
+After all geometry has been projected into the CDDT, each bin is sorted to facilitate later indexing.
+Not only does the direct construction of the CDDT greatly reduce the amount of memory required to store the LUT, it also reduces data structure precomputation time since ray casting is not necessary.
+
+[TODO add an example of scene rotation, saw tooth wave, and zero point projection]
+
+\subsection{CDDT Construction Algorithm}
+
+\subsection{CDDT Query Algorithm}
+
+{\color{red}DONE WITH SECOND EDITING PASS UP TO THIS POINT}
 
@@
 
-\subsection{Rotational Symmetry}
+\subsection{Further Optimizations}
 
@@
 All benchmarks were performed on a desktop computer on a Intel Core i5-4590 CPU @ 3.30GHz with 16GB DDR3 ram @ 1333MHz, running Ubuntu 14.04.
-
-% \begin{figure}[h]
-% \img{grid_trial2_basement_cropped.png}
-% \caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method.
-Basement map.
-X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
-% \label{violin:basement:grid}
-% \end{figure}
 
@@
 CDDT and PCDDT exploiting radial symmetry while resolving the sensor model.}
-\label{violin:basement:random}
+\label{violin:basement:particle}
 \end{figure}
@@
 
+\bibitem{acceleratedraymarching} Zuiderveld, Karel, Koning, Anton, and Viergever, Max, “Acceleration of ray-casting using 3D distance transforms”, Proceedings of the SPIE – Visualization in Biomedical Computing 1992, Vol.1808, pp.324-335, 1992.
+
 \bibitem{sphere:source} Pharr, M., and R.

### 2017-02-27 includes/RangeLib.h

@@ -197,6 +197,12 @@ namespace ranges {
 			for (int y = 0; y < height; ++y) {
 				for (int x = 0; x < width; ++x) {
 					unsigned idx = 4 * y * width + 4 * x;
+
+					// if (trace_grid[x][y]) {
+					// 	image[idx + 0] = 255;
+					// 	image[idx + 1] = 255;
+					// 	image[idx + 2] = 255;
+					// }
 					
 					image[idx + 2] = 255;
 					image[idx + 1] = 255;
@@ -205,7 +211,7 @@ namespace ranges {
 					if (trace_grid[x][y]) {
 						image[idx + 0] = 0;
 						image[idx + 1] = 0;
-						image[idx + 2] = 255;
+						image[idx + 2] = 200;
 					}
 
 					if (grid[x][y]) {
@@ -1526,6 +1532,8 @@ namespace benchmark {
 			std::cout << " -avg time per ray: " << ( t_accum / (float) num_samples) << " sec" << std::endl;
 			std::cout << " -rays cast: " << num_samples << std::endl;
 		}
+
+		ranges::OMap *getMap() {return range.getMap(); }
 
 	protected:
 		range_T range;

### 2017-02-27 main.cpp

@@ -35,11 +35,15 @@ DEFINE_string(cddt_save_path, "",
 DEFINE_string(log_path, "",
 	"Where to store high fidelity logs, does not save if not specified.");
 
-DEFINE_string(which_benchmark, "random",
+DEFINE_string(which_benchmark, "",
 	"Which benchmark to run, one of:\n  random\n  grid\n");
 
+DEFINE_string(query, "",
+	"Query point x,y,theta to ray cast from. example: --query=0,0,3.14");
+
+DEFINE_string(trace_path, "", "Path to output trace map of memory access pattern. Works for Bresenham's Line or Ray Marching.");
+
 #define MAX_DISTANCE 500
-#define NUM_RAYS 500
 #define THETA_DISC 108
 #define MB (1024.0*1024.0)
 // #ifdef BASEPATH
@@ -50,7 +54,7 @@ DEFINE_string(which_benchmark, "random",
 #define GRID_STEP 10
 #define GRID_RAYS 40
 #define GRID_SAMPLES 1
-#define RANDOM_SAMPLES 200000
+#define RANDOM_SAMPLES 100
 
 using namespace ranges;
 using namespace benchmark;
@@ -106,6 +110,15 @@ int main(int argc, char *argv[])
 
 	if (map.error()) return 1;
 
+	float query_y, query_x, query_t;
+	if (!FLAGS_trace_path.empty()) {
+		std::vector<std::string> query = utils::split(FLAGS_query, ',');
+		query_x = std::stof(query[0]);
+		query_y = std::stof(query[1]);
+		query_t = std::stof(query[2]);
+	}
+	
+
 	std::vector<std::string> methods = utils::split(FLAGS_method, ',');
 
 	if (utils::has("BresenhamsLine", methods) || utils::has("bl", methods)) {
@@ -129,6 +142,21 @@ int main(int argc, char *argv[])
 			mark.grid_sample(GRID_STEP, GRID_RAYS, GRID_SAMPLES);
 		else if (FLAGS_which_benchmark == "random")
 			mark.random_sample(RANDOM_SAMPLES);
+
+		if (!FLAGS_query.empty()) {
+			std::cout << "...querying pose:" << FLAGS_query << std::endl;
+			std::cout << "...   range: " << bl.calc_range(query_x,query_y,query_t) << std::endl;
+		}
+
[…]

### 2017-02-28 includes/RangeLib.h
paper changes

@@ -308,6 +308,15 @@ namespace ranges {
 
 		DistanceTransform() : width(0), height(0) {}
 
+		DistanceTransform(int w, int h) : width(w), height(h) {
+			// allocate space in the vectors
+			for (int i = 0; i < width; ++i) {
+				std::vector<float> y_axis;
+				for (int q = 0; q < height; ++q) y_axis.push_back(1.0);
+				grid.push_back(y_axis);
+			}
+		}
+
 		// computes the distance transform of a given OMap
 		DistanceTransform(OMap *map) {
 			width = map->width;
@@ -346,12 +355,23 @@ namespace ranges {
 			lodepng::State state; //optionally customize this one
 			char image[width * height * 4];
 
+			float scale = 0;
+			for (int y = 0; y < height; ++y) {
+				for (int x = 0; x < width; ++x) {
+					scale = std::max(grid[x][y], scale);
+				}
+			}
+			scale *= 1.0 / 255.0;
 			for (int y = 0; y < height; ++y) {
 				for (int x = 0; x < width; ++x) {
 					unsigned idx = 4 * y * width + 4 * x;
-					image[idx + 2] = std::min(255, (int)grid[x][y]);
-					image[idx + 1] = std::min(255, (int)grid[x][y]);
-					image[idx + 0] = std::min(255, (int)grid[x][y]);
+					// std::cout << (int)(grid[x][y] / scale) << " " << grid[x][y] / scale << std::endl;
+					// image[idx + 2] = std::min(255, (int)grid[x][y]);
+					// image[idx + 1] = std::min(255, (int)grid[x][y]);
+					// image[idx + 0] = std::min(255, (int)grid[x][y]);
+					image[idx + 2] = (int)(grid[x][y] / scale);
+					image[idx + 1] = (int)(grid[x][y] / scale);
+					image[idx + 0] = (int)(grid[x][y] / scale);
 					image[idx + 3] = 255;
 				}
 			}
@@ -1433,6 +1453,24 @@ namespace ranges {
 			#else
 			return giant_lut[(int)x][(int)y][discretize_theta(heading)];
 			#endif
+		}
+
+		DistanceTransform *get_slice(float theta) {
+			int width = giant_lut.size();
+			int height = giant_lut[0].size();
+			DistanceTransform *slice = new DistanceTransform(width, height);
+
+			int dtheta = discretize_theta(theta);
+
+			for (int x = 0; x < width; ++x) {
[…]

### 2017-02-28 main.cpp

@@ -43,8 +43,11 @@ DEFINE_string(query, "",
 
 DEFINE_string(trace_path, "", "Path to output trace map of memory access pattern. Works for Bresenham's Line or Ray Marching.");
 
+DEFINE_string(lut_slice_path, "", "Path to output a slice of the LUT.");
+DEFINE_string(lut_slice_theta, "1.57", "Which LUT slice to output");
+
 #define MAX_DISTANCE 500
-#define THETA_DISC 108
+#define THETA_DISC 16
 #define MB (1024.0*1024.0)
 // #ifdef BASEPATH
 // #define FLAGS_log_path BASEPATH "/tmp/basement/"
@@ -290,6 +293,12 @@ int main(int argc, char *argv[])
 			mark.grid_sample(GRID_STEP, GRID_RAYS, GRID_SAMPLES);
 		else if (FLAGS_which_benchmark == "random")
 			mark.random_sample(RANDOM_SAMPLES);
+
+		if (!FLAGS_lut_slice_path.empty()) {
+			std::cout << "...saving LUT slice theta="  << FLAGS_lut_slice_theta << " to: " << FLAGS_lut_slice_path << std::endl;
+			float theta = std::stof(FLAGS_lut_slice_theta);
+			glt.get_slice(theta)->save(FLAGS_lut_slice_path);
+		}
 		if (DO_LOG) {
 			save_log(tlog, FLAGS_log_path+"/glt.csv");
 		}

### 2017-02-28 visualize.py

@@ -5,6 +5,7 @@ from yaml import CLoader as Loader, CDumper as Dumper
 import ujson
 import itertools
 import argparse
+import scipy.misc
 
 # dump = yaml.dump(dummy_data, fh, encoding='utf-8', default_flow_style=False, Dumper=Dumper)
 # data = yaml.load(fh, Loader=Loader)
@@ -41,7 +42,7 @@ class CDDTSlice(object):
         max_zero = max(map(max, non_empty_zeros))
         return [int(np.ceil(max_zero - min_zero))+1,len(self.zeros)]
 
-    def make_ddt(self, saw_tooth=True):
+    def make_ddt(self, saw_tooth=True, reversed_dir=False):
         non_empty_zeros = filter(lambda x: len(x) > 0, self.zeros)
         if len(non_empty_zeros) == 0:
             print "Empty slice, nothing to visualize"
@@ -51,24 +52,45 @@ class CDDTSlice(object):
         min_zero = min(map(min, non_empty_zeros))
         max_zero = max(map(max, non_empty_zeros))
         height = int(np.ceil(max_zero - min_zero))+1
-        ddt = np.zeros((height,len(self.zeros)))
+
+        grid_height = len(self.zeros)
+        # ddt = np.zeros((height,len(self.zeros)))
+        ddt = np.zeros((grid_height,len(self.zeros)))
+        offset = int((grid_height - height) / 2.0)
 
         for x in xrange(len(self.zeros)):
             for zp in self.zeros[x]:
-                y = int(zp - min_zero)
+                y = int(zp - min_zero+offset)
                 ddt[y,x] = 1
 
         if saw_tooth:
             for x in xrange(len(self.zeros)):
-                last = -1
-                for y in xrange(height):
-                    if ddt[y,x] == 1:
-                        last = 0
-                        ddt[y,x] = last
-                    elif last >= 0:
-                        last = last + 1
-                        ddt[y,x] = last
+                if reversed_dir:
+                    last = -1
+                    for y in reversed(xrange(grid_height)):
+                        if ddt[y,x] == 1:
+                            last = 0
+                            ddt[y,x] = last
+                        elif last >= 0:
[…]

### 2017-02-28 paper/main.tex
another major revision

@@
 \usetikzlibrary{shapes.geometric, arrows}
+\usepackage{algorithm}
+% \usepackage[]{algorithm2e}
+\usepackage{algorithmic}
 
@@
 \newcommand{\imgFrame}[1]{\begin{center}\frame{\includegraphics[width=\columnwidth]{{#1}}}\end{center}}
+\newcommand{\INDSTATE}[1][1]{\STATE\hspace{#1\algorithmicindent}}
 
@@
 \title{\LARGE \bf
-Compressed Directional Distance Transform for Fast 2D Ray Casting
+Fast Approximate 2D Ray Casting for Accelerated Localization
+% Compressed Directional Distance Transform for Fast 2D Ray Casting
 }
@@
 \pagestyle{empty}
-
-\skcomment{}{I would consider putting the word ``localization'' in the title.}
 
@@
 In robots with range sensors such as LiDAR or Sonar, ray casting is frequently used to compare sensor readings with the ground truth distance between the hypothesis pose and obstacles in a map.
-Ray casting itself is a complex operation, and reweighting any single particle may require tens of ray casts.
+Ray casting itself is a complex operation, and a single evaluation of the sensor model may require tens of ray casts.
 Many effective particle filters maintain thousands of particles, updating each particle tens of times per second.
@@
 The contributions of this work are the following.
-Firstly, we develop a problem formulation and provide details of our method.
+Firstly, we develop a problem formulation and provide details and analysis of our method.
 Secondly, we implement the proposed algorithm and compare its performance to alternate two-dimensional ray casting methods.
@@
 
-Due to varied handling of edge cases and ambiguities in the problem of determining distance between any point and the nearest obstacle in a particular direction, most 2D ray casting algorithms do not provide exactly consistent results for every query.
-However, on average error between results from each of the methods discussed is small.
-
-Bresenham's line algorithm \cite{bresenham} incrementally determines the set of pixels that approximate the trajectory of query ray starting from the query point $(x,y)$ and progressing in the $\theta$ direction one pixel at a time.
-The algorithm terminates once the nearest occupied pixel is discovered, and the euclidean distance between that occupied pixel and the query $(x,y)$ is reported.
+Bresenham's line algorithm \cite{bresenham} incrementally determines the set of pixels that approximate the trajectory of query ray starting from the query point $(x,y)_{query}$ and progressing in the $\theta_{query}$ direction one pixel at a time.
+The algorithm terminates once the nearest occupied pixel is discovered, and the euclidean distance between that occupied pixel and $(x,y)_{query}$ is reported.
 This algorithm is widely implemented in particle filters due to its simplicity and ability to operate on a dynamic map.
@@
 The primary difference is that ray marching makes larger steps along the query ray, thereby avoiding unnecessary memory reads.
-Beginning at the query point $(x,y)$ the ray marching algorithm proceeds in the theta direction, stepping along the line by the minimum distance between each query point and the nearest obstacle.
+Beginning at $(x,y)_{query}$ the ray marching algorithm proceeds in the $\theta_{query}$ direction, stepping along the line by the minimum distance between each query point and the nearest obstacle.
 The algorithm terminates when the query point coincides with an obstacle in the map.
@@
 
-Ray marching is on average much faster than Bresenham’s line, but edge cases exist in which the performance is nearly equivalent.
+Ray marching is on average much faster than Bresenham’s line, but edge cases exist in which the theoretically asymptotic runtime is equivalent.
 As noted in \cite{acceleratedraymarching} the traversal speed of rays rapidly decreases as sampled positions approach obstacles.
@@
 In the authors’ implementation ray casting is used to populate the LUT since it has low initialization time and is significantly faster than Bresenham’s Line.
-Theoretically this approach has constant query runtime, though actual performance is dependent on access patterns as CPU caching effects are significant in practice.
+Theoretically the LUT approach has constant query runtime, though actual performance is dependent on access patterns as CPU caching effects are significant in practice.
 
 State space discretization implies approximate results, since intermediate states must be rounded onto the discrete grid.
-While the effect of rounding query position, it can be significant for $\theta$.
+While the effect of rounding query position is small, rounding $\theta_{query}$ may have significant effects.
 As the ray moves further away from the query point along the theta direction, angular discretization error accumulates.
-For queries $(x,y,\theta)$ discretized into $(x_d,y_d,\theta_d)$, the distance between the end of the ray $(x_d,y_d,Θ_d)$ and its projection onto the line implied by $(x,y,\theta)$ becomes large as the length of the ray increases.
+For queries $(x,y,\theta)_{query}$ discretized into $\lfloor(x,y,\theta)\rceil$, the distance between the end of the ray $\lfloor(x,y,\theta)\rceil$ and its projection onto the line implied by $(x,y,\theta)_{query}$ becomes large as the length of the ray increases.
 Although the error induced by discretization may be unacceptable for computer graphics applications, it is generally acceptable for probabilistic algorithms such as the particle filter since error is expected and directly modeled.
 
-Rather than simply rounding to the nearest grid state, one method to improve accuracy of queries which lie between discrete states is to query the neighboring states and interpolate between them.
-Since we are evaluating the performance of these methods for use in a particle filter, we do not perform interpolation since the increased computation reduces the number of particles that may be maintained in real time, and therefore any increase in accuracy with respect to the stored map is effectively negated.
+More generally, due to varied handling of edge cases and ambiguities in the problem of determining distance between any point and the nearest obstacle in a particular direction, most 2D ray casting algorithms do not provide exactly consistent results for every query.
+However, on average error between results from each of the methods discussed is small.
+Any error that does exist is dominated by error inherent in distance sensors.
+
+Rather than simply rounding to the nearest grid state, one may improve accuracy of queries which lie between discrete states by querying the neighboring states and interpolating the results.
+Since we are evaluating the performance of these methods for use in a particle filter, we do not perform interpolation since the increased computation reduces the number of particles that may be maintained in real time, thereby effectively negating any potential increase in accuracy with respect to the stored map.
 
@@
 We define the problem of ray casting in occupancy grids as follows.
-We assume a known occupancy grid map in which occupied cells have value one and unoccupied cells have value zero.
-Given a query pose $(x,y,\theta)_{query}$ in map space, the ray cast operation finds the nearest occupied pixel $(x,y)_{collide}$ which lies on the ray starting at the position $(x,y)_{query}$ pointing in the $\theta_{query}$ direction.
-The ray cast operation returns the euclidean distance $d_{ray}$ between $(x,y)_{query}$ and $(x,y)_{collide}$.
+We assume a known occupancy grid map in which occupied cells have value 1, and unoccupied cells have value 0.
+Given a query pose $(x,y,\theta)_{query}$ in map space, the ray cast operation finds the nearest occupied pixel $(x,y)_{collide}$ which lies on the ray starting at the position $(x,y)_{query}$ pointing in the $\theta_{query}$ direction, and returns the euclidean distance $d_{ray}$ between $(x,y)_{query}$ and $(x,y)_{collide}$.
 
@@
 We refer to $(x,y,\theta)_{query}$ as the query pose, and $(x,y)_{query}$ as the query point.
-Additionally, denote the discretized query pose as $\lfloor(x,y,\theta)_{query}\rceil$.
+We denote the discretized query pose as $\lfloor(x,y,\theta)_{query}\rceil$.
 A $\theta$ slice through the LUT refers to all table entries for that particular $\lfloor\theta\rceil$.
+The number of discrete $\lfloor\theta\rceil$ values is denoted $\theta_{discretization}$.
+Fig.
+\ref{coordinate_system} demonstrates our chosen coordinate system.
 
@@
 Our data structure is designed to compress this redundancy while still allowing for fast queries in near constant time.
-We accomplish this though through what we refer to as a Compressed Directional Distance Transform (CDDT) described here.
+We accomplish this though through what we refer to as a Compressed Directional Distance Transform described here.
 
@@
 
-In contrast, to compute the DDT, the obstacles in the map are rotated about the origin by $-\lfloor\theta\rceil$ and ray casting is implicitly performed in the $\theta=0$ direction to populate the table.
-Then, to index the table for a query $(x,y,\theta)_{query}$ one discretizes to $\lfloor(x,y,\theta)_{query}\rceil$, rotates $\lfloor(x,y)_{query}\rceil$ about the origin, and uses the rotated coordinates $\lfloor(x,y,\theta)_{query}\rceil_{rot}$ to index into the DDT.
-Thus, the same operation is computed in both the DDT and the LUT, but while one changes the ray cast direction to populate the LUT, one rotates scene geometry and ray casts in a constant direction to populate the DDT.
-While it is true that the rotation of the scene geometry introduces small errors as compared to ray casting for each cell, as previously discussed small errors do not significantly impact particle filter localization performance due to the use of a probabilistic error model and are dominated by errors inherently present in distance sensors.
+\begin{figure}[h]
+\img{ddt_lut_slice_comparison_map.png}
+\caption{Comparison between a DDT and a LUT slice for the same value of theta. Each row in the DDT is characterized by a sawtooth function.}
+\label{ddt_lut_comparison}
+\end{figure}
+
+In contrast, to compute the DDT, the obstacles in the map are rotated about the origin by $-\lfloor\theta\rceil$ and ray casting is implicitly performed in the $\theta=0$ direction, as demonstrated by Fig.
+\ref{ddt_lut_comparison}.
+Then, to index the table for a query $(x,y,\theta)_{query}$ one discretizes to $\lfloor(x,y,\theta)_{query}\rceil$, rotates $\lfloor(x,y)_{query}\rceil$ about the origin, and uses the rotated coordinates $\lfloor(x,y,\theta)_{query}\rceil_{rot}$ to (implicitly) index into the DDT.
+Thus, roughly the same operation is computed in both the DDT and the LUT, but while one changes the ray cast direction to populate the LUT, one rotates scene geometry and ray casts in a constant direction to populate the DDT.
+
+We define $P_{DDT_{\theta}}$ as the three by three transformation matrix which projects vectors into the coordinate space of the DDT in the $\theta$ direction.
+While it is true that the transformation of the scene geometry introduces small errors as compared to ray casting for each cell, as previously discussed, small errors do not significantly impact particle filter localization performance.
 
@@
 
-The conversion from a DDT to a CDDT slice is performed by storing the $x$ coordinate of each zero point in the coordinate space of the DDT for every discrete row.
-At query time, the sawtooth function (which is the distance to the nearest obstacle in the query direction) encoded in the DDT may be quickly recovered by performing binary search for nearest zero point in the correct row of the CDDT slice.
+\begin{figure}[h]
+\img{sawtooth_to_cddt.png}
+\caption{Demonstration of compression from a sawtooth function to a list of zero points. Consecutively valued zero points exist where walls are greater than one pixel thick.}
+\label{sawtooth_to_cddt}
+\end{figure}
+
+The conversion from a DDT to a CDDT slice is performed by storing the $x$ coordinate of each zero point in the coordinate space of the DDT for every discrete row as shown in Fig.
+\ref{sawtooth_to_cddt}.
+At query time, the sawtooth function (i.e.
+exactly the distance to the nearest obstacle in the query direction) encoded in the DDT may be quickly recovered by performing binary search for nearest zero point in the correct row of the CDDT slice.
 We refer to the list of zero points for a single row as a CDDT bin.
@@
 
-For performance, it is not necessary to compute the full DDT, but rather the CDDT may be directly constructed by projecting each obstacle into the coordinate space of the DDT for every $\lfloor\theta\rceil$ and sorting its $x$ coordinate in the CDDT bin corresponding to its $y$ coordinate and rotation.
+For performance, it is not necessary to compute the full DDT, but rather the CDDT may be directly constructed by projecting each obstacle into the coordinate space of the DDT for every $\lfloor\theta\rceil$ and sorting its $x$ coordinate in the CDDT bin corresponding to its $y$ coordinate and $\lfloor\theta\rceil$.
 After all geometry has been projected into the CDDT, each bin is sorted to facilitate later indexing.
@@
 
-[TODO add an example of scene rotation, saw tooth wave, and zero point projection]
-
-\subsection{CDDT Construction Algorithm}
-
-\subsection{CDDT Query Algorithm}
-
-{\color{red}DONE WITH SECOND EDITING PASS UP TO THIS POINT}
-
-Once the CDDT is constructed, querying the data structure to resolve ray casts involves
-
-\begin{enumerate}
-    \item Projecting the query point into the coordinate space of the slice which corresponds to the discretized theta
-    \item Finding the nearest zero point which has a y-coordinate greater than or equal to the query point in the projected coordinate space
-    \item Return the distance between the query point and the nearest zero point in the projected coordinate space, which is simply the difference between y-coordinates
-\end{enumerate}
-
-Binary search is theoretically optimal for finding zero points near to query states, in our implementation we conditionally use linear search when the number of zero points is below a certain constant (we use 64) as it has a superior memory access pattern in such cases.
 While conceptually simple, implementing the CDDT data structure construction and traversal routines requires careful consideration in order to capture all edge cases and to minimize unnecessary computation.
@@
 
-\skcomment{}{You can put all the text after this point into one subsection, which you can call ``Further optimizations'' or something like that.}
+\subsection{CDDT Construction Algorithm}
 
-\subsection{Further Optimizations}
+\begin{center}
+\line(1,0){240}
+\end{center}
+\begin{algorithmic}
+% \STATE {\color{gray}\# Finds all occupied pixels bordering unoccupied pixels}
+\STATE $edge\_map\gets map-morphological\_erosion(map)$
+\STATE Initialize $\theta_{discretization}$ empty CDDT slices
+\FOR {$\theta \in \{\lfloor\theta\rceil\}$} 
+        \FOR{each occupied pixel $(x,y)\in edge\_map$}
+            \STATE $(x,y)_{DDT_{\theta}} = P_{DDT_{\theta}}*\begin{pmatrix}
+                        x \\
+                        y \\
+                        1
+             \end{pmatrix}$
+             \FOR {each CDDT bin overlapping with $y_{DDT_{\theta}}$}
+                \STATE bin.append($x_{DDT_{\theta}}$)
+             \ENDFOR
+        \ENDFOR
+        \FOR{each CDDT bin}
+            \STATE bin = sort(bin)
+        \ENDFOR
+\ENDFOR
+\end{algorithmic}
+\begin{center}
+\line(1,0){240}
+\end{center}
 
-A side effect of extracting the zero points from each column of the DDT is the introduction of rotational symmetry.
-For any given theta, the zero points accumulated by projecting scene geometry into the requisite coordinate space are mirror images of those accumulated by projecting scene geometry into the theta+pi direction.
-Therefore, one need only compute the CDDT for the range of discrete thetas from $[0,\pi)$ and the DDT for the range $[0,2\pi)$ may be inferred, resulting in factor of two further reduction in memory usage.
+\subsection{CDDT Query Algorithm}
+
+\begin{center}
+\line(1,0){240}
+\end{center}
+\begin{algorithmic}
+\STATE \textbf{function ray\_cast($(x,y,\theta)_{q})$}
+\INDSTATE[1] $(x,y)_{DDT_{\theta_{q}}} = P_{DDT_{\theta_{q}}}*\begin{pmatrix}
+                        x_{q} \\
+                        y_{q} \\
+                        1
+             \end{pmatrix}$
+\INDSTATE[1] bin $\gets$ zero points in row $y_{DDT_{\theta_{q}}}$ of CDDT slice $\theta_{q}$
+\INDSTATE[1] $x_{collide}$ = smallest element $x_{collide} > x_{DDT_{\theta_q}}\in$ bin
+% find nearest greate element in CDDT bin correspond to y_ddt
+\INDSTATE[1] \textbf{return} abs($x_{DDT_{\theta_q}} - x_{collide}$)
+\STATE \textbf{end function}\end{algorithmic}
+\begin{center}
+\line(1,0){240}
+\end{center}
+
+In our implementation we switch between linear and binary search to find $x_{collide}$ in a given CDDT bin depending on the number of zero points in that bin.
+While binary search is theoretically optimal for sorted arrays, in practice linear search has a superior memory access pattern and is therefore faster in small searches.
+
+\subsection{Further Optimizations} \label{optimizations}
+
+A side effect of extracting zero points from each row of the DDT is the introduction of rotational symmetry.
+Not only can a row of the DDT in the $\theta$ direction be reconstructed from the zero points, but also a row of the DDT in the $\theta+\pi$ direction as in Fig.
+\ref{bidirectional_zero_points}.
+Therefore, one only need compute CDDT slices for $0\leq\lfloor\theta\rceil<\pi$ and the DDT for $0\leq\lfloor\theta\rceil<2\pi$ may be inferred, resulting in a factor of further two reduction in memory usage.
+
+\begin{figure}[h]
+\img{bidirectional_zero_points.png}
+\caption{Demonstration of reconstructing rows of two DDT slices from a single set of zero points.}
+\label{bidirectional_zero_points}
+\end{figure}
 
 In addition to the aforementioned reduction in memory usage, rotational symmetry may be exploited in scenarios where ray casts are performed radially around a single point.
-While traversing the data structure to resolve a ray cast in a particular $(x,y,\theta)$, the ray cast for $(x,y,\theta+\pi)$ may be resolved with a single additional memory read.
-Once the index i of the nearest zero point in the theta direction $z_{\theta}$ is discovered, the index of the nearest zero point in the opposite direction is simply i-1.
-For example, in robots with laser scanners sweeping angles larger than 180°, this symmetry can be used to reduce the number of data structure traversals required to compute the sensor model by up to a factor of two.
+While traversing the data structure to resolve a ray cast query $(x,y,\theta)$, the ray cast for $(x,y,\theta+\pi)$ may be resolved with a single additional memory read.
+Once a search algorithm is used to discover the index $i$ of $x_{collide_{\theta}}$ in the CDDT, the index of $x_{collide_{\theta+\pi}}$ is simply $i-1$ as in Fig.
+\ref{bidirectional_zero_points}.
+For example, this symmetry can be used in robots with laser scanners sweeping angles larger than $180^{\circ}$ to reduce the number of data structure traversals required to compute the sensor model by up to a factor of two.
 
-\subsection{Incremental Update}
+% \subsection{Incremental Update}
 
 While the authors did not implement the necessary changes, we conjecture that the CDDT algorithm could be modified to allow incremental updates to the map.
-In contrast, any change to the map would introduce inconsistencies in a precomputed LUT requiring a full re-computation of the data structure.
-Ray casting every state in a three dimensional grid is highly computational, ruling out the possibility of using a precomputed LUT with dynamic maps such as in SLAM algorithms.
-Similarly, ray marching operates on a euclidean distance transform which cannot be easily updated in an incremental fashion.
-While Bresenham’s line algorithm is the slowest option evaluated in terms of ray cast performance, it has the benefit that it operates directly on the occupancy grid map and can therefore be used with dynamic maps.
-For this reason, Bresenham’s line is one of the most commonly used algorithms for ray casting in occupancy grids.
+In both ray marching and the LUT approach, any change in the map introduces inconsistencies to the acceleration data structure requiring a full re-initialization.
+As demonstrated by our experiments, initializing the three dimensional LUT is a highly computational process.
+Similarly, computing a euclidean distance transform for ray marching can be expensive.
+Therefore neither ray casting nor the LUT can be used with dynamic maps such as in SLAM algorithms.
+Despite being the slowest option evaluated, Bresenham's Line algorithm is one of the most commonly used techniques since it operates directly on the grid map without an acceleration data structure.
 
-To allow incremental update of the CDDT, the sorted vector data structure used in our implementation should be replaced with an alternative which allows for $O(logn)$ insert and delete, such as a randomized skip list.
-To aid in node deletion, pointers could be maintained from each grid element to the associated leaf node in the randomized skip lists with a constant factor increase in memory usage.
-
-In this case, the cost of inserting or deleting an occupied grid cell would be \skcomment{$O(theta\_discretization*\log(number of lut bin elements))$.}{I think this is your plan already, but just to note: Please make sure to use symbols for these and introduce what the symbols mean.
-} The number of elements in each lut bin tends to be small and is bounded, so this becomes $O(theta\_discretization)$ which is likely not prohibitive for real-time performance.
+To allow incremental update of the CDDT, the sorted vector data structure used in our implementation to store zero points should be replaced with an alternative which allows for $O(\log{n})$ insertion and deletion, such as a randomized skip list.
+To aid in zero point deletion, pointers could be maintained correlating each cell in the occupancy grid with its associated zero points for each CDDT slice with a constant factor increase in memory usage.
+In this case, the cost of toggling the state of a cell in the occupancy grid would be $O(\theta_{discretization}\log{n})$ where $n$ is the number of elements in each associated CDDT bin.
+For most occupancy grid maps of real environments, the value of $n$ is small, and in any case is bounded by the size of the map, so $\log{n}$ is bounded by a small constant factor.
+Therefore, the cost of update with becomes $O(\theta_{discretization})$ which is likely not prohibitive for real-time performance in dynamic maps.
 
 Skip list traversal is likely to be less efficient than binary search on sorted vectors due to worse cache characteristics.
-However, the authors expect it to still be significantly faster than performing Bresenham’s line, and therefore recommend that this modification be the subject of future research.
+However, the authors expect it to still be faster than performing Bresenham’s Line algorithm, and therefore recommend that this modification be the subject of future research.
 
-\subsection{Pruning}
+% \subsection{Pruning}
+\begin{figure}[h]
+\begin{center}\includegraphics[width=6.5cm]{edge_map.png}\end{center}
+\caption{Example map left, morphological edge map right.}
+\label{edge_map}
+\end{figure}
 
 By removing entries in the CDDT which can never possibly result in a ray collision, it is possible to further reduce the memory footprint of the data structure.
-\skcomment{Consider a 3x3 block of obstacles in an otherwise empty map.}{I think this deserves a figure, just to allow the reader to quickly read your paper.} The center obstacle will never be the nearest neighbor in any ray casting query, because any such query would first intersect with one of the 8 surrounding obstacles.
+Consider a 3x3 block of obstacles in an otherwise empty map as in Fig.
+\ref{edge_map}.
+The center obstacle will never be the nearest neighbor in any ray casting query, because any such query would first intersect with one of the 8 surrounding obstacles.
 To exploit this, one can take the morphological edge map of the occupancy grid map prior to CDDT generation without loss of generality.
-This can result in significant memory usage and data structure construction time reductions in dense maps where the number of obstacles adjacent to empty cells is small compared to the total number of obstacles.
-Prior to performing ray casting on the CDDT data structure, we check the occupancy grid to ensure that the query point itself is not an obstacle and return a distance of zero if so in order to prevent incorrect results when ray casting from the middle of obstacles which may be removed by the morphological operation.
+Prior to performing ray casting on the CDDT data structure, we check the occupancy grid to ensure that the query point itself is not an obstacle and return a distance of zero if so in order to prevent incorrect results when ray casting from the middle of obstacles removed during the morphological operation.
+In dense maps, this results in a significant reduction in both memory usage and construction time.
 
-\begin{figure}[h]
-\img{slice0_pcddt_basement.png}
-\caption{The same slice of the DDT as \ref{cddt:slice0}.
-Notice that the number of elements projected into each CDDT bin is less peaky since the pruning operation removes most elements of long lines of obstacles.
-Also note the higher compression factor of 415 for this slice, increased from 267 in the unpruned CDDT.}
-\label{pcddt:slice0}
-\end{figure}
+% \begin{figure}[h]
+% \img{slice0_pcddt_basement.png}
+% \caption{A slice of the PCDDT visualized as the implied DDT. Blue line demonstrates the num}
+% \label{pcddt:slice0}
+% \end{figure}
 
-\skcomment{Additionally, consider a line of obstacles aligned along the Y-axis.}{While you are at it, put a figure for this one too?
-} Every element in this line will be projected into a single list of zero points for the theta=0 slice of the CDDT.
+Additionally, consider a line of obstacles aligned along the X-axis.
+Every element in this line will be projected into a single zero point bin in the $\theta=0$ CDDT slice.
 However, the middle elements of the line will never result in a collision.
-Any ray cast from adjacent states in the $\theta=0$ direction will return early in the occupancy grid check, and any ray cast from outside the wall will collide with one of the edge-most obstacles in that line.
-Therefore in the theta=0 slice, one may discard the zero points corresponding to the middle elements without introducing error.
-This form of optimization is simple to compute in the cardinal directions, but non-trivial for arbitrary theta not aligned with an axis.
-Rather than attempting to analytically determine which obstacles may be discarded, it is simpler to prune the data structure by ray casting from every possible state (as in the LUT data structure) and discard any zero point which is unused.
+Any ray cast from points on the line of obstacles will return early in the occupancy grid check, and any ray cast from non-overlapping points co-linear in the $\theta$ or $\theta+\pi$ directions will intersect one of the edge-most obstacles in the line.
+Therefore in the $\theta=0$ CDDT slice, one may discard the zero points corresponding to the middle elements without introducing error.
+This form of optimization is simple to compute in the cardinal directions, but non-trivial for arbitrary $\lfloor\theta\rceil$ not aligned with an axis.
+Rather than analytically determining which obstacles may be discarded, it is simpler to prune the data structure by ray casting from every possible state, discarding any unused zero point.
 
-Pruning does also increase precomputation time rendering it incompatible with incremental update, however the reduction of memory usage is worthwhile for static maps.
-In addition to memory reduction, we find that pruning slightly improves runtime performance, likely as a result of improved caching characteristics.
+Pruning does increase pre-computation time, rendering it incompatible with incremental update.
+However, the reduction of memory usage is worthwhile for static maps.
+In addition to memory reduction, we find that pruning slightly improves runtime performance, likely as a result of improved caching characteristics and the reduced number of zero points.
 
-\skcomment{}{Do you think you can have a subsection called ``analysis'' which has a few propositions/theorems that say things like Memory requirement is $O(NM)$ and running time is $O(1)$?
+\section{Analysis}
+
+% \skcomment{}{Do you think you can have a subsection called ``analysis'' which has a few propositions/theorems that say things like Memory requirement is $O(NM)$ and running time is $O(1)$?
 This may be hard given the deadline, but osmething to think about.
@@
 
-\section{EXPERIMENTS}
+The memory usage of the CDDT data structure is $O(n*\theta_{discretization}+|map|)$ where $n$ is the number of occupied pixels in the edge map.
+The occupancy grid map must be stored to check for overlaps between $(x,y)_{query}$ and obstacles prior to searching CDDT bins.
+For each edge pixel, a total of $\theta_{discretization}$ float values are stored in the CDDT bins.
+Due to the fact that we must sort each bin after CDDT construction, pre-computation time is at worst $O(n*\theta_{discretization} + S^2*\theta_{discretization}*log(S))$ for the same definition of n, where $S=norm(\text{map width},\text{map height})$.
+In practice it is closer to $O(n*\theta_{discretization} + S*\theta_{discretization})$ since each CDDT has a small number of elements on average, as evidenced by the high demonstrated compression ratio.
+
+The pruning operation described in \ref{optimizations} reduces memory requirement, with a computational cost of $O(|map|*\theta_{discretization}*O(calc\_range)_{CDDT})$.
+The precise impact of pruning is scene dependent and difficult to analyze in the general case.
+
+The ray cast procedure has three general steps: projection into CDDT coordinate space, the search for nearby zero points, and the computation of distance given the nearest zero point.
+The first and last steps are simple arithmetic, and therefore are theoretically constant time.
+The second step involves searching for a value in a sorted array of zero points.
+As previously discussed in \ref{optimizations}, the number of zero points in each CDDT bin tends to be small and is bounded in map size.
+Thus, at worst the search operation requires $\log{(norm(\text{map width},\text{map height}))}$ which is a constant value for a fixed map.
+Therefore, for a given map, our algorithm provides constant time performance.
+
+\section{Experiments}
 
@@
 All benchmarks were performed on a desktop computer on a Intel Core i5-4590 CPU @ 3.30GHz with 16GB DDR3 ram @ 1333MHz, running Ubuntu 14.04.
+
+\begin{figure}[h!]
+\begin{center}
+\begin{tabular}{ | m{8em} | m{2.3cm}| m{1.5cm} | } 
+\hline
+ \multicolumn{3}{|c|}{Basement Map, $\theta_{discretization}$: 108} \\
+ \hline
+ Method & Memory Usage & Init. Time \\
+ \hline
+ Bresenham's Line & 1.37 MB & 0.006 sec  \\
+ Ray Marching & 5.49 MB & 0.16 sec  \\
+ CDDT & 6.34 MB & 0.067 sec  \\
+ PCDDT & 4.07 MB & 2.2 sec  \\
+ Lookup Table & 296.63 MB & 15.3 sec  \\
+\hline \hline
+% \end{center}
+% % \caption{The construction time and memory footprint of each method for the Basement map.}
+% \label{table:basement:init}
+% \end{figure}
+% \begin{figure}[ht]
+% \begin{center}
+% \begin{tabular}{ | m{8em} | m{2.3cm}| m{1.5cm} | } 
+ \multicolumn{3}{|c|}{Synthetic Map, $\theta_{discretization}$: 108} \\
+ \hline
+ Method & Memory Usage & Init. Time \\
+ \hline
+ Bresenham's Line & 1 MB & 0.004 sec  \\
+ Ray Marching  & 4 MB & 0.13 sec  \\
+ CDDT & 2.71 MB & 0.03 sec  \\
+ PCDDT & 1.66 MB & 0.74 sec  \\
+ Lookup Table & 216 MB & 9.1 sec  \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Construction time and memory footprint of each method.}
+\label{table:synthetic:init}
+\end{figure}
 
@@
 
-\subsection{Synthetic Benchmarks}
+We evaluate algorithm performance in two synthetic benchmarks, using two different maps.
+The so called Synthetic map was created with photoshop, whereas the basement map was created via a SLAM algorithm on the RACECAR platform {\color{red} [cite this?]}.
+The first benchmark computes a ray cast for each point in a uniformly spaced grid over the three dimensional state space.
+The second benchmark performs a large number of ray casts for randomly generated states.
 
@@
 
-We evaluate algorithm performance in two synthetic benchmarks, using two different maps.
-The so called Synthetic map was created with photoshop, whereas the basement map was created via a SLAM algorithm on the RACECAR platform [cite this?].
-The first benchmark computes a ray cast for each point in a uniformly spaced grid over the three dimensional state space.
-The second benchmark performs a large number of ray casts for randomly generated states.
-
-\begin{figure}[ht]
-\begin{center}
-\begin{tabular}{ | m{8em} | m{2.3cm}| m{1.5cm} | } 
-\hline
- \multicolumn{3}{|c|}{Basement Map, $\theta$ discretization: 108} \\
- \hline
- Method & Memory Usage & Init. Time \\
- \hline
- Bresenham's Line & 1.37 MB & 0.006 sec  \\
- Ray Marching & 5.49 MB & 0.16 sec  \\
- CDDT & 6.34 MB & 0.067 sec  \\
- PCDDT & 4.07 MB & 2.2 sec  \\
- Lookup Table & 296.63 MB & 15.3 sec  \\
-\hline
-\end{tabular}
-\end{center}
-\caption{The construction time and memory footprint of each method for the Basement map.}
-\label{table:basement:init}
-\end{figure}
-
-\begin{figure}[ht]
-\begin{center}
-\begin{tabular}{ | m{8em} | m{2.3cm}| m{1.5cm} | } 
-\hline
- \multicolumn{3}{|c|}{Synthetic Map, $\theta$ discretization: 108} \\
- \hline
- Method & Memory Usage & Init. Time \\
- \hline
- Bresenham's Line & 1 MB & 0.004 sec  \\
- Ray Marching  & 4 MB & 0.13 sec  \\
- CDDT & 2.71 MB & 0.03 sec  \\
- PCDDT & 1.66 MB & 0.74 sec  \\
- Lookup Table & 216 MB & 9.1 sec  \\
-\hline
-\end{tabular}
-\end{center}
-\caption{The construction time and memory footprint of each method for the Synthetic map.}
-\label{table:synthetic:init}
-\end{figure}
-
-\subsection{Particle Filter Experiments}
-
 \begin{figure}[h]
@@
 To simulate real world operating conditions, we have implemented a particle filter localization algorithm using a beam mode sensor model.
-We provide statistics about the ray cast performance of each algorithm while being used to compute the sensor model, as well as end to end statistics on the number of particles able to be maintained in real-time.
+We provide statistics about the ray cast performance of each algorithm while being used to compute the sensor model.
 While care was taken in implementing the particle filter, there are undoubtedly many optimizations which could further improve end-to-end performance.
-Thus, these benchmarks should not be considered an upper bound on performance and should be interpreted comparatively.
+Thus, we do not present end-to-end particle filter statistics.
 
-Our sensor model is designed for the Hokuyo UST-10LX lidar scanner used aboard the RACECAR platform [cite this?], which features a 270° field of view.
-Since this FOV is in excess of 180° we exploit radial symmetry discussed in [radial symmetry section] to simultaneously ray cast in the theta and $\theta+180°$ direction when possible.
-As demonstrated in the below figure, this optimization reduces the number of data structure traversals required by a third as compared to the other methods evaluated, while still resolving the same number of ray casts.
-As is standard in such applications, we downsample the resolution of the laser scanner to reduce the number of ray casts per sensor model evaluations and to make the probability distribution over the state space less peaked.
+Our sensor model is designed for the Hokuyo UST-10LX lidar scanner used aboard the RACECAR platform, which features a $270^{\circ}$ field of view.
+Since this FOV is in excess of $180^{\circ}$ we exploit radial symmetry discussed in the subsection \ref{optimizations} to simultaneously ray cast in the theta and $\theta+\pi$ direction when possible.
+This optimization reduces the number of data structure traversals required by a third while still resolving the same number of ray casts.
+As is standard in such applications, we down-sample the laser scanner resolution to reduce the number of ray casts per sensor model evaluation, and to make the probability distribution over the state space less peaked.
 
-\section{CONCLUSIONS}
+\section{Conclusions}
 
 This work demonstrates that the proposed CDDT algorithm may be used in mobile robotics to accelerate sensor model computation when localizing in a two dimensional occupancy grid map.
-While the precomputed LUT approach is generally 1.1 to 1.7 times faster than the proposed algorithm, the memory footprint of the proposed data structure is significantly smaller which may be desirable for resource constrained systems.
+While the precomputed LUT approach is generally 1.1 to 1.7 times faster than the proposed algorithm, the memory footprint of the proposed data structure is approximately two orders of magnitude smaller, which may be desirable for resource constrained systems.
 Other than the LUT approach, CDDT is significantly faster than the other methods evaluated.
@@
 
+While our experiments reveal a multi-modal distribution of completion times for the CDDT algorithm, it does not exhibit a significant long tail distribution, confirming our constant time analysis.
+We suspect the various modes in completion time are due to short circuit cases in our implementation, as well as caching and kernel interrupt effects.
+
 In future work we aim to further develop our approach to allow for incremental map updates without requiring a full reconstruction of the underlying acceleration data structure, a feature which currently only the slowest evaluated method supports.
-
-\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
+\addtolength{\textheight}{-2cm}   % This command serves to balance the column lengths
                                   % on the last page of the document manually. It shortens
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{APPENDIX}
+\section*{Appendix}
 
 \begin{figure}[h!]
-\img{synthetic_cropped.png}
+% \img{synthetic.png}
+\begin{center}\includegraphics[width=4.5cm]{synthetic_border.png}\end{center}
+\setlength{\belowcaptionskip}{-10pt}
 \caption{Synthetic map occupancy grid. 1024x1024}
@@
 
-\begin{figure}[h!]
+\begin{figure}[ht]
 \begin{center}
 \newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}}
-\begin{tabular}{ | C{1.15cm} | C{1.25cm}| C{1.25cm} | C{1.25cm} | C{1.25cm}|  } 
+\begin{tabular}{ | C{1.15cm} | C{1.25cm}| C{1.25cm} | C{1.24cm} | C{1.23cm}|  } 
 \hline
@@
 
-\bibitem{localization} Sebastian Thrun, Dieter Fox, Wolfram Burgard and Frank Dellaert. “Robust Monte Carlo Localization for Mobile Robots.” Artificial Intelligence Journal. 2001
-
 \bibitem{bresenham} J. Bresenham, “Algorithm for Computer Control of a Digital Plotter,” IBM Systems Journal, vol. 4, no. 1, pp. 25-30, 1965
@@
 \bibitem{raymarching} K. Perlin and E. M. Hoffert, “Hypertexture.” Computer Graphics, vol 23, no. 3, pp. 297-306, 1989.
+
+\bibitem{localization} Sebastian Thrun, Dieter Fox, Wolfram Burgard and Frank Dellaert. “Robust Monte Carlo Localization for Mobile Robots.” Artificial Intelligence Journal. 2001
 

### 2017-03-02 paper/main.tex
added submission version of paper

@@
 \usepackage{algorithm}
-% \usepackage[]{algorithm2e}
 \usepackage{algorithmic}
@@
 \newcommand{\img}[1]{\begin{center}\includegraphics[width=\columnwidth]{{#1}}\end{center}}
-
 \newcommand{\imgFrame}[1]{\begin{center}\frame{\includegraphics[width=\columnwidth]{{#1}}}\end{center}}
 \newcommand{\INDSTATE}[1][1]{\STATE\hspace{#1\algorithmicindent}}
-
 \newcommand{\gridScale}{1.25}
@@
 %: Margin note commands
-\usepackage[usenames,dvipsnames]{xcolor}
-\setlength{\marginparwidth}{0.5in}
-\let\oldmarginpar\marginpar
-\renewcommand\marginpar[1]{\-\oldmarginpar[\raggedleft\footnotesize #1]%
-{\raggedright\footnotesize #1}}
-\newcommand{\skcomment}[2]{{\color{red}#1} \marginpar{\color{blue}\tiny #2 (SK)}}
-
-% \begin{center}\frame{\includegraphics[width=\columnwidth]{synthetic.png}}\end{center}
+% \usepackage[usenames,dvipsnames]{xcolor}
+% \setlength{\marginparwidth}{0.5in}
+% \let\oldmarginpar\marginpar
+% \renewcommand\marginpar[1]{\-\oldmarginpar[\raggedleft\footnotesize #1]%
+% {\raggedright\footnotesize #1}}
+% \newcommand{\skcomment}[2]{{\color{red}#1} \marginpar{\color{blue}\tiny #2 (SK)}}
 
@@
 
-Lorem Ipsum is simply dummy text of the printing and typesetting industry.
-Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.
-It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.
-It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.
+Localization is an essential component for autonomous robots.
+A well-established localization approach combines ray casting with a particle filter, leading to a computationally expensive algorithm that is difficult to run on resource-constrained mobile robots.
+We present a novel data structure called the Compressed Directional Distance Transform for accelerating ray casting in two dimensional occupancy grid maps.
+Our approach has near constant time ray casting performance for a fixed size map, in contrast with other methods which exhibit long tail query time distributions.
+Our experimental results show that the proposed algorithm approximates the performance characteristics of reading from a three dimensional lookup table of ray cast solutions while requiring two orders of magnitude less memory and precomputation.
+
+% Ray casting is a computationally expensive and commonly used operation in particle filter localization algorithms onboard mobile robots.
+We present a novel data structure called the Compressed Directional Distance Transform for accelerating ray casting in two dimensional occupancy grid maps.
+Our approach has near constant time ray casting performance for a fixed size map, in contrast with other methods which exhibit long tail query time distributions.
+Our experimental results show that the proposed algorithm approximates the performance characteristics of reading from a three dimensional lookup table of ray cast solutions while requiring two orders of magnitude less memory and precomputation.
 
@@
 Both algorithms work by iteratively checking points in the map starting at the query point and moving in the ray direction until an obstacle is discovered.
-This process may require hundreds of memory reads per ray cast depending on the distance to the nearest obstacle, and does not provide constant time performance.
+This process does not provide constant performance because the number of memory reads depends on the distance to the nearest obstacle.
+% This process requires many memory reads, and does not provide constant time performance.
 
-To combat the computational challenges of ray casting while localizing in a two-dimensional map, Thrun et al.
+To combat the computational challenges of ray casting while localizing in a two-dimensional map, Fox et al.
 \cite{localization} suggest the use of a large three-dimensional lookup table (LUT) to store expected ranges for each discrete $(x,y,\theta)$ state.
@@
 Finally, we discuss potential extensions to the CDDT algorithm which could be the basis of future research.
+
 We observe two orders of magnitude improvement in memory requirements, with little sacrifice in computation time, when compared to the lookup table methods.
@@
 
-Bresenham's line algorithm \cite{bresenham} incrementally determines the set of pixels that approximate the trajectory of query ray starting from the query point $(x,y)_{query}$ and progressing in the $\theta_{query}$ direction one pixel at a time.
+Bresenham's line algorithm \cite{bresenham} incrementally determines the set of pixels that approximate the trajectory of a query ray starting from the query point $(x,y)_{query}$ and progressing in the $\theta_{query}$ direction one pixel at a time.
 The algorithm terminates once the nearest occupied pixel is discovered, and the euclidean distance between that occupied pixel and $(x,y)_{query}$ is reported.
@@
 The primary disadvantage is that it is slow, potentially requiring hundreds of memory accesses for a single ray cast.
-While actual performance is highly environment dependent, Bresenham’s Line algorithm is linear in map size in the worst case.
+While average performance is highly environment dependent, Bresenham’s Line algorithm is linear in map size in the worst case.
 
@@
 
-As previously described, a common acceleration technique for two dimensional ray casting is to precompute the ray distances for every state in a discrete grid and store the results in a three dimensional LUT.
+As previously described, a common acceleration technique for two dimensional ray casting is to precompute the ray distances for every state in a discrete grid and store the results in a three dimensional LUT for later reference.
 Any ray casting technique may be used to compute the table.
-In the authors’ implementation ray casting is used to populate the LUT since it has low initialization time and is significantly faster than Bresenham’s Line.
+In the authors’ implementation ray marching is used for its fast performance compared to Bresenham's Line and its high accuracy.
 Theoretically the LUT approach has constant query runtime, though actual performance is dependent on access patterns as CPU caching effects are significant in practice.
@@
 State space discretization implies approximate results, since intermediate states must be rounded onto the discrete grid.
-While the effect of rounding query position is small, rounding $\theta_{query}$ may have significant effects.
-As the ray moves further away from the query point along the theta direction, angular discretization error accumulates.
+While the effect of rounding query position is small, rounding $\theta_{query}$ may have significant effects since angular roundoff error accumulates with ray length.
 For queries $(x,y,\theta)_{query}$ discretized into $\lfloor(x,y,\theta)\rceil$, the distance between the end of the ray $\lfloor(x,y,\theta)\rceil$ and its projection onto the line implied by $(x,y,\theta)_{query}$ becomes large as the length of the ray increases.
-Although the error induced by discretization may be unacceptable for computer graphics applications, it is generally acceptable for probabilistic algorithms such as the particle filter since error is expected and directly modeled.
+Although the error induced by discretization may be unacceptable for computer graphics applications, it is generally acceptable for probabilistic algorithms such as the particle filter since error is anticipated and directly modeled.
 
@@
 Rather than simply rounding to the nearest grid state, one may improve accuracy of queries which lie between discrete states by querying the neighboring states and interpolating the results.
-Since we are evaluating the performance of these methods for use in a particle filter, we do not perform interpolation since the increased computation reduces the number of particles that may be maintained in real time, thereby effectively negating any potential increase in accuracy with respect to the stored map.
+Interpolation requires greater than one memory read and a small amount of arithmetic, and is therefore slower on average than rounding to the nearest grid state.
+Since we are evaluating the performance of these methods in the context of particle filter localization, the number of particles maintained is more important than ray cast accuracy so we do not perform interpolation.
 
@@
 
-\begin{figure}[h]
-\img{coord_system.png}
+\begin{figure}[h!]
+% \img{}
+\begin{center}\includegraphics[width=7.6cm]{coord_system.png}\end{center}
 \caption{Occupancy grid map coordinate system}
@@
 
-Although the three dimensional table used to store precomputed ray cast solutions for a discrete state space is inherently large, it is highly compressible.
+Although the three dimensional table used to store precomputed ray cast solutions in a discrete state space is inherently large, it is highly compressible.
 This is most apparent in the cardinal directions, in which adjacent values along a particular dimension of the table increase by exactly one unit of distance for unobstructed positions as in Fig.
@@
 \end{center}
-\caption{Occupancy grid (left) and associated LUT for $\theta=0$ (right).}
+\caption{Occupancy grid (left) and associated LUT slice for $\theta=0$ (right).}
 \label{cardinal_example}
@@
 The euclidean distance transform of an occupancy grid map stores the distance to the nearest obstacle in any direction $\theta$ for each possible $\lfloor(x,y)\rceil$ in the map.
-In contrast, a directional distance transform (DDT) stores the distance to the nearest obstacle in a particular direction $\lfloor\theta\rceil$ for every $\lfloor(x,y)\rceil$.
+In contrast, what we call a directional distance transform (DDT) stores the distance to the nearest obstacle in a particular direction $\lfloor\theta\rceil$ for every $\lfloor(x,y)\rceil$.
 The key difference between a single $\theta$ slice of the LUT and a directional distance transform for the same $\theta$ is the way in which it is computed and indexed.
@@
 \begin{figure}[h]
+\vspace{4pt}
 \img{ddt_lut_slice_comparison_map.png}
@@
 \ref{ddt_lut_comparison}.
-Then, to index the table for a query $(x,y,\theta)_{query}$ one discretizes to $\lfloor(x,y,\theta)_{query}\rceil$, rotates $\lfloor(x,y)_{query}\rceil$ about the origin, and uses the rotated coordinates $\lfloor(x,y,\theta)_{query}\rceil_{rot}$ to (implicitly) index into the DDT.
+Then, to index the table for a query $(x,y,\theta)_{query}$, one discretizes to $\lfloor(x,y,\theta)_{query}\rceil$, rotates $\lfloor(x,y)_{query}\rceil$ about the origin, and uses the rotated coordinates $\lfloor(x,y,\theta)_{query}\rceil_{rot}$ to (implicitly) index into the DDT.
 Thus, roughly the same operation is computed in both the DDT and the LUT, but while one changes the ray cast direction to populate the LUT, one rotates scene geometry and ray casts in a constant direction to populate the DDT.
@@
 \img{sawtooth_to_cddt.png}
-\caption{Demonstration of compression from a sawtooth function to a list of zero points. Consecutively valued zero points exist where walls are greater than one pixel thick.}
+\caption{Demonstration of compression from a sawtooth function to a list of zero points. Consecutively valued zero points exist projected scene geometry spans greater than one unit along the x axis.}
 \label{sawtooth_to_cddt}
@@
 
-The conversion from a DDT to a CDDT slice is performed by storing the $x$ coordinate of each zero point in the coordinate space of the DDT for every discrete row as shown in Fig.
+% The conversion from a DDT to a CDDT slice is done by storing the $x$ coordinate of each zero point in the coordinate space of the DDT for every discrete row as shown in Fig. \ref{sawtooth_to_cddt}.
+
+The conversion from DDT to CDDT slice is done by storing the $x$ coordinates of each zero point for every row of the DDT as demonstrated for a single row in Fig.
 \ref{sawtooth_to_cddt}.
@@
 
-For performance, it is not necessary to compute the full DDT, but rather the CDDT may be directly constructed by projecting each obstacle into the coordinate space of the DDT for every $\lfloor\theta\rceil$ and sorting its $x$ coordinate in the CDDT bin corresponding to its $y$ coordinate and $\lfloor\theta\rceil$.
+For performance, it is not necessary to compute or store the full DDT, but rather the CDDT may be directly constructed by projecting each obstacle into the coordinate space of the DDT for every $\lfloor\theta\rceil$ and storing its projected $x$ coordinate in the CDDT bin corresponding to its $y$ coordinate and $\lfloor\theta\rceil$.
 After all geometry has been projected into the CDDT, each bin is sorted to facilitate later indexing.
-Not only does the direct construction of the CDDT greatly reduce the amount of memory required to store the LUT, it also reduces data structure precomputation time since ray casting is not necessary.
+Not only does the direct construction of the CDDT greatly reduce the amount of memory required to store a lookup table, it also reduces data structure precomputation time since ray casting every discrete state is not necessary.
 
@@
 In this sense, it is more complex to implement than the alternatives considered, however there are many opportunities for optimization which yield real-world speed up.
-To ease this burden, we provide our implementation as well as Python wrappers with an Apache 2.0 license.
+To ease this burden, we provide our implementation\footnote{https://github.com/kctess5/range\_libc} as well as Python wrappers under the Apache 2.0 license.
 
@@
 \ref{bidirectional_zero_points}.
-Therefore, one only need compute CDDT slices for $0\leq\lfloor\theta\rceil<\pi$ and the DDT for $0\leq\lfloor\theta\rceil<2\pi$ may be inferred, resulting in a factor of further two reduction in memory usage.
+Therefore, one only need compute and store CDDT slices for $0\leq\lfloor\theta\rceil<\pi$ and the DDT for $0\leq\lfloor\theta\rceil<2\pi$ may be inferred by reversing the search direction, resulting in a factor of two reduction in memory usage.
 
@@
 
-% \subsection{Incremental Update}
+With small modifications, another factor of two reduction in CDDT size could be attained by storing zero points as 16 bit integers rather than 32 bit floats with little precision loss.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Incremental Update %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Similarly, computing a euclidean distance transform for ray marching can be expensive.
-Therefore neither ray casting nor the LUT can be used with dynamic maps such as in SLAM algorithms.
+Therefore neither ray marching nor the LUT can be easily used with dynamic maps such as in SLAM algorithms.
 Despite being the slowest option evaluated, Bresenham's Line algorithm is one of the most commonly used techniques since it operates directly on the grid map without an acceleration data structure.
@@
 
-% \subsection{Pruning}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Pruning %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{figure}[h]
@@
 To exploit this, one can take the morphological edge map of the occupancy grid map prior to CDDT generation without loss of generality.
-Prior to performing ray casting on the CDDT data structure, we check the occupancy grid to ensure that the query point itself is not an obstacle and return a distance of zero if so in order to prevent incorrect results when ray casting from the middle of obstacles removed during the morphological operation.
+During ray casting on the CDDT data structure, we check the occupancy grid to ensure that the query point itself is not an obstacle and return a distance of zero if so in order to prevent incorrect results when ray casting from the middle of obstacles removed during the morphological operation.
 In dense maps, this results in a significant reduction in both memory usage and construction time.
-
-% \begin{figure}[h]
-% \img{slice0_pcddt_basement.png}
-% \caption{A slice of the PCDDT visualized as the implied DDT. Blue line demonstrates the num}
-% \label{pcddt:slice0}
-% \end{figure}
 
@@
 However, the middle elements of the line will never result in a collision.
-Any ray cast from points on the line of obstacles will return early in the occupancy grid check, and any ray cast from non-overlapping points co-linear in the $\theta$ or $\theta+\pi$ directions will intersect one of the edge-most obstacles in the line.
+Any ray cast from points on the line of obstacles will return early in the occupancy grid check, and any ray cast from non-overlapping points co-linear in the $\theta$ or $\theta+\pi$ directions will intersect either the first or last obstacle in the line.
 Therefore in the $\theta=0$ CDDT slice, one may discard the zero points corresponding to the middle elements without introducing error.
@@
 In addition to memory reduction, we find that pruning slightly improves runtime performance, likely as a result of improved caching characteristics and the reduced number of zero points.
+We refer to the pruned datastructure as PCDDT.
 
 \section{Analysis}
-
-% \skcomment{}{Do you think you can have a subsection called ``analysis'' which has a few propositions/theorems that say things like Memory requirement is $O(NM)$ and running time is $O(1)$?
-This may be hard given the deadline, but osmething to think about.
-}
 
@@
 Due to the fact that we must sort each bin after CDDT construction, pre-computation time is at worst $O(n*\theta_{discretization} + S^2*\theta_{discretization}*log(S))$ for the same definition of n, where $S=norm(\text{map width},\text{map height})$.
-In practice it is closer to $O(n*\theta_{discretization} + S*\theta_{discretization})$ since each CDDT has a small number of elements on average, as evidenced by the high demonstrated compression ratio.
+In practice it is closer to $O(n*\theta_{discretization} + S*\theta_{discretization})$ since each CDDT bin has a small number of elements on average, as evidenced by the high demonstrated compression ratio.
 
@@
 As previously discussed in \ref{optimizations}, the number of zero points in each CDDT bin tends to be small and is bounded in map size.
-Thus, at worst the search operation requires $\log{(norm(\text{map width},\text{map height}))}$ which is a constant value for a fixed map.
-Therefore, for a given map, our algorithm provides constant time performance.
+Thus, at worst the search operation requires $\log{(norm(\text{map width},\text{map height}))}$ which is a small constant value for a fixed map.
+Therefore, for a given map, our algorithm provides $O(1)$ query performance.
 
@@
 We have implemented the proposed algorithm in the C++ programming language, as well as Bresenham’s Line, ray marching, and the LUT approach for comparison.
-Our source code is available for use and analysis [TODO add a link to github], and Python wrappers are also provided for easier usage.
+Our source code$^1$ is available for use and analysis, and Python wrappers are also provided for easier usage.
 All benchmarks were performed on a desktop computer on a Intel Core i5-4590 CPU @ 3.30GHz with 16GB DDR3 ram @ 1333MHz, running Ubuntu 14.04.
@@
 \hline \hline
-% \end{center}
-% % \caption{The construction time and memory footprint of each method for the Basement map.}
-% \label{table:basement:init}
-% \end{figure}
-% \begin{figure}[ht]
-% \begin{center}
-% \begin{tabular}{ | m{8em} | m{2.3cm}| m{1.5cm} | } 
  \multicolumn{3}{|c|}{Synthetic Map, $\theta_{discretization}$: 108} \\
@@
 \end{center}
-\caption{Construction time and memory footprint of each method.}
+\caption{Construction time and memory footprint of each method. Ranges stored in lookup table with 16 bit integers, initialized with ray marching.}
 \label{table:synthetic:init}
@@
 \begin{figure}[ht]
+\vspace{4pt}
 \img{basement_grid_random_cropped.png}
@@
 We evaluate algorithm performance in two synthetic benchmarks, using two different maps.
-The so called Synthetic map was created with photoshop, whereas the basement map was created via a SLAM algorithm on the RACECAR platform {\color{red} [cite this?]}.
+The so called Synthetic map (Fig.
+\ref{synthetic_map}) was created with photoshop, whereas the basement map (Fig.
+\ref{sawtooth_to_cddt}) was created via a SLAM algorithm on the RACECAR platform\footnote{http://racecar.mit.edu} while navigating the Stata basement loop.
 The first benchmark computes a ray cast for each point in a uniformly spaced grid over the three dimensional state space.
@@
 
-% \begin{figure}[h]
-% \img{grid_trial2_basement_cropped.png}
-% \caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method.
-Basement map.
-X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
-% \label{violin:basement:grid}
-% \end{figure}
-
-% \begin{figure}[h]
-% \img{grid_trial2_synthetic_cropped.png}
-% \caption{Violin plots demonstrating histogram of completion time over a large number of queries for each ray cast method.
-Synthetic map.
-X axis shows time in milliseconds, and Y axis shows the number of queries that completed after that amount of time.}
-% \label{violin:basement:grid}
-% \end{figure}
-
-% \begin{figure}[h]
-% \img{random_trial2_basement_cropped.png}
-% \caption{Violin plots for a large number of randomly sampled queries. Basement map. The random access pattern degrades the caching performance variably for each method.}
-% \label{violin:basement:random}
-% \end{figure}
-
-% \begin{figure}[h]
-% \img{random_trial2_synthetic_cropped.png}
-% \caption{Violin plots for a large number of randomly sampled queries. Synthetic map. The random access pattern degrades the caching performance variably for each method.}
-% \label{violin:basement:random}
-% \end{figure}
-
 \begin{figure}[h]
+\vspace{2pt}
 \img{particle_trial2_basement_cropped.png}
@@
 While our experiments reveal a multi-modal distribution of completion times for the CDDT algorithm, it does not exhibit a significant long tail distribution, confirming our constant time analysis.
-We suspect the various modes in completion time are due to short circuit cases in our implementation, as well as caching and kernel interrupt effects.
+We suspect the various modes in completion time are due to short circuit cases in our implementation, as well as caching effects.
 
 In future work we aim to further develop our approach to allow for incremental map updates without requiring a full reconstruction of the underlying acceleration data structure, a feature which currently only the slowest evaluated method supports.
-\addtolength{\textheight}{-2cm}   % This command serves to balance the column lengths
+\addtolength{\textheight}{0cm}   % This command serves to balance the column lengths
                                   % on the last page of the document manually. It shortens
@@
 \caption{Synthetic map occupancy grid. 1024x1024}
-\label{violin:basement:grid}
+\label{synthetic_map}
 \end{figure}
 
-\begin{figure}[ht]
+\begin{figure}[h]
 \begin{center}
+\vspace{4pt}
 \newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}}
@@
 
-\begin{figure}[ht]
+\begin{figure}[h]
 \begin{center}
@@
 \end{tabular}
-
 \caption{All times listed in seconds, speedup relative to Bresenham's Line}
@@
 
-% \begin{figure}[h!]
-% \begin{center}
-% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
-% \hline
-% \multicolumn{5}{|c|}{Basement Map, Random Sampling Ray Cast} \\
-% \hline
-% Method & Mean & Median & IQR & Speedup \\
-% \hline
-% BL & 9.66e-07 & 1.05e-06 & 1.08e-06 & 1 \\
-% RM & 2.12e-07 & 1.68e-07 & 1.64e-07 & 4.56 \\
-% CDDT & 1.58e-07 & 1.49e-07 & 9.1e-08 & 6.13 \\
-% PCDDT & 1.41e-07 & 1.32e-07 & 6.5e-08 & 6.83 \\
-% LUT & 1.89e-07 & 1.7e-07 & 2.1e-08 & 5.10 \\
-% \hline
-% \end{tabular}
-% \end{center}
-% \caption{Statistics per ray cast query for random sampling synthetic map.}
-% \label{table:basement:init}
-% \end{figure}
-% \begin{figure}[h!]
-% \begin{center}
-% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.25cm}|  } 
-% \hline
-% \multicolumn{5}{|c|}{Basement Map, Grid Sampling Ray Cast} \\
-% \hline
-% Method & Mean & Median & IQR & Speedup \\
-% \hline
-% BL & 8.29e-07 & 9.24e-07 & 9.53e-07 & 1 \\
-% RM & 1.65e-07 & 1.26e-07 & 1.34e-07 & 5.02 \\
-% CDDT & 7.69e-08 & 7.2e-08 & 1.6e-08 & 10.78 \\
-% PCDDT & 7.32e-08 & 7e-08 & 1.4e-08 & 11.33 \\
-% LUT & 6.13e-08 & 4.3e-08 & 4.6e-08 & 13.53 \\
-% \hline
-% \end{tabular}
-% \end{center}
-% \caption{Statistics per ray cast query for grid sampling synthetic map.}
-% \label{table:basement:init}
-% \end{figure}
-% \begin{figure}[h!]
-% \begin{center}
-% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.15cm}|  } 
-% \hline
-% \multicolumn{5}{|c|}{Basement Map, Particle Filter Ray Cast} \\
-% \hline
-% Method & Mean & Median & IQR & Speedup \\
-% \hline
-% BL & 3.89e-07 & 2.67e-07 & 2.31e-07 & 1 \\
-% RM & 1.56e-07 & 1.18e-07 & 9.4e-08 & 2.49 \\
-% CDDT & 7.67e-08 & 7.8e-08 & 5.7e-08 & 5.07 \\
-% PCDDT & 7.22e-08 & 7.4e-08 & 5.4e-08 & 5.39 \\
-% LUT & 4.35e-08 & 4.1e-08 & 5e-09 & 8.94 \\
-% \hline
-% \end{tabular}
-% \end{center}
-% \caption{Statistics per ray cast query for random sampling basement map.}
-% \label{table:basement:init}
-% \end{figure}
-% \begin{figure}[h!]
-% \begin{center}
-% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.15cm}|  } 
-% \hline
-% \multicolumn{5}{|c|}{Synthetic Map, Random Sampling Ray Cast} \\
-% \hline
-% Method & Mean & Median & IQR & Speedup \\
-% \hline
-% BL & 1.19e-06 & 1.41e-06 & 7.71e-07 & 1 \\
-% RM & 1.52e-07 & 1.25e-07 & 1.05e-07 & 7.81 \\
-% CDDT & 1.24e-07 & 1.05e-07 & 5.3e-08 & 9.59 \\
-% PCDDT & 1.19e-07 & 1.01e-07 & 5e-08 & 10.02 \\ 
-% LUT & 1.82e-07 & 1.68e-07 & 1.4e-08 & 6.55 \\
-% \hline
-% \end{tabular}
-% \end{center}
-% \caption{The construction time and memory footprint of each method for the Basement map.}
-% \label{table:basement:init}
-% \end{figure}
-% \begin{figure}[h!]
-% \begin{center}
-% \begin{tabular}{ | m{1.05cm} | m{1.3cm}| m{1.6cm} | m{1.25cm} | m{1.15cm}|  } 
-% \hline
-% \multicolumn{5}{|c|}{Synthetic Map, Grid Sampling Ray Cast} \\
-% \hline
-% Method & Mean & Median & IQR & Speedup \\
-% \hline
-% BL & 1.03e-06 & 1.20e-06 & 6.79e-07 & 1 \\ 
-% RM & 1.27e-07 & 1.03e-07 & 1.06e-07 & 8.06 \\
-% CDDT & 7.02e-08 & 6.8e-08 & 1e-08 & 14.63 \\
-% PCDDT & 6.94e-08 & 6.8e-08 & 9e-09 & 14.80 \\
-% LUT & 6.33e-08 & 4.2e-08 & 4.6e-08 & 16.21 \\
-% \hline
-% \end{tabular}
-% \end{center}
-% \caption{The construction time and memory footprint of each method for the Basement map.}
-% \label{table:basement:init}
-% \end{figure}
-
 \begin{thebibliography}{99}
 
-\bibitem{bresenham} J. Bresenham, “Algorithm for Computer Control of a Digital Plotter,” IBM Systems Journal, vol. 4, no. 1, pp. 25-30, 1965
+ \bibitem{bresenham} J. Bresenham. ``Algorithm for Computer Control of a Digital Plotter,'' \textit{IBM Systems Journal}, vol. 4, no. 1, pp. 25-30, 1965.
 
-\bibitem{raymarching} K. Perlin and E. M. Hoffert, “Hypertexture.” Computer Graphics, vol 23, no. 3, pp. 297-306, 1989.
+\bibitem{raymarching} K. Perlin and E. M. Hoffert. ``Hypertexture,'' \textit{Computer Graphics}, vol 23, no. 3, pp. 297-306, 1989.
 
-\bibitem{localization} Sebastian Thrun, Dieter Fox, Wolfram Burgard and Frank Dellaert. “Robust Monte Carlo Localization for Mobile Robots.” Artificial Intelligence Journal. 2001
+\bibitem{localization} D.
+Fox, W.
+Burgard, and S.
+Thrun.
+``Markov localization for mobile robots in dynamic environments,'' \textit{Journal of Artificial Intelligence Research}, vol.
+11, pp.
+391–427, 1999.
 
-\bibitem{acceleratedraymarching} Zuiderveld, Karel, Koning, Anton, and Viergever, Max, “Acceleration of ray-casting using 3D distance transforms”, Proceedings of the SPIE – Visualization in Biomedical Computing 1992, Vol.1808, pp.324-335, 1992.
+\bibitem{sphere:source} M.
+Pharr, and R.
+Fernando.
+``Chapter 8.
+Per-Pixel Displacement Mapping with Distance Functions'' in \textit{GPU gems 2: Programming techniques for high-performance graphics and general-purpose computation}, 3rd ed.
+United States: Addison-Wesley Educational Publishers, 2005.
 
-\bibitem{sphere:source} Pharr, M., and R.
-Fernando.
-GPU Gems 2: Programming Techniques for High-Performance Graphics and General-Purpose Computation: Addison Wesley Professional.
-2005.
-http://http.developer.nvidia.com/GPUGems2/gpugems2\_chapter08.html
+\bibitem{acceleratedraymarching} K.
+Zuiderveld, A.
+Koning, and M.
+Viergever.
+``Acceleration of ray-casting using 3D distance transforms,'' \textit{Proceedings of the SPIE}, vol.
+1808, pp.
+324-335, 1992.
 

### 2017-03-02 includes/RangeLib.h

@@ -68,6 +68,9 @@ Useful Links: https://github.com/MRPT/mrpt/blob/4137046479222f3a71b5c00aee1d5fa8
 #define _USE_LRU_CACHE 0
 #define _LRU_CACHE_SIZE 1000000
 
+// not implemented yet -> use 16 bit integers to store zero points
+#define _CDDT_SHORT_DATATYPE 1
+
 #define _GIANT_LUT_SHORT_DATATYPE 1
 
 // slow unoptimized version 
@@ -1364,6 +1367,7 @@ namespace ranges {
 			limits_div_max = std::numeric_limits<uint16_t>::max() / max_range;
 			#endif
 			RayMarching seed_cast = RayMarching(m, mr);
+			// CDDTCast seed_cast = CDDTCast(m, mr, td);
 
 			for (int x = 0; x < m.width; ++x) {
 				std::vector<std::vector<lut_t> > lut_slice;

### 2017-03-02 main.cpp

@@ -47,7 +47,7 @@ DEFINE_string(lut_slice_path, "", "Path to output a slice of the LUT.");
 DEFINE_string(lut_slice_theta, "1.57", "Which LUT slice to output");
 
 #define MAX_DISTANCE 500
-#define THETA_DISC 16
+#define THETA_DISC 108
 #define MB (1024.0*1024.0)
 // #ifdef BASEPATH
 // #define FLAGS_log_path BASEPATH "/tmp/basement/"
@@ -283,6 +283,7 @@ int main(int argc, char *argv[])
 
 		Benchmark<GiantLUTCast> mark = Benchmark<GiantLUTCast>(glt);
 		std::cout << "...lut size (MB): " << glt.memory() / MB << std::endl;
+		std::cout << "...construction time: " << construction_dur.count() << std::endl;
 		std::cout << "...Running grid benchmark" << std::endl;
 		if (DO_LOG) {
 			tlog.str("");

### 2017-03-05 includes/CudaRangeLib.h
removed the dopevector thing, added GPU ray marching

@@ -1,0 +1,17 @@
+#include <vector>
+
+
+class RayMarchingCUDA
+{
+public:
+	RayMarchingCUDA(std::vector<std::vector<float> > grid, int w, int h, float mr);
+	~RayMarchingCUDA();
+	void calc_range_many(float *ins, float *outs, int num_casts);
+private:
+	float *d_ins;
+	float *d_outs;
+	float *d_distMap;
+	int width;
+	int height;
+	float max_range;
+};

### 2017-03-05 includes/RangeLib.h

@@ -29,10 +29,8 @@ Useful Links: https://github.com/MRPT/mrpt/blob/4137046479222f3a71b5c00aee1d5fa8
 #define RANGE_LIB_H
 
 #include "vendor/lodepng/lodepng.h"
-// #include "vendor/dt/dt.h"
-
+#include "vendor/distance_transform.h"
 #include "includes/RangeUtils.h"
-#include "vendor/distance_transform/include/distance_transform/distance_transform.hpp"
 
 #include <stdio.h>      /* printf */
 #include <cstdlib>
@@ -102,6 +100,15 @@ Useful Links: https://github.com/MRPT/mrpt/blob/4137046479222f3a71b5c00aee1d5fa8
 #define J2 J1 J1
 #define J3 J1 J1 J1
 #define J4 J2 J2
+
+#ifdef USE_CUDA
+	// #define CHUNK_SIZE (64*256)
+	// #define NUM_THREADS (256)
+	#include "includes/CudaRangeLib.h"
+	// #include <cuda.h>
+	// #include <cuda_runtime_api.h>
+#endif
+
 
 namespace ranges {
 	struct OMap
@@ -249,10 +256,10 @@ namespace ranges {
 				for (int x = 0; x < width; ++x) {
 					unsigned idx = 4 * y * width + 4 * x;
 					
-					image[idx + 2] = 255;
-					image[idx + 1] = 255;
-					image[idx + 0] = 255;
-					image[idx + 3] = 255;
+					image[idx + 2] = (char)255;
+					image[idx + 1] = (char)255;
+					image[idx + 0] = (char)255;
+					image[idx + 3] = (char)255;
 
 					if (grid[x][y]) {
 						image[idx + 0] = 0;
@@ -270,7 +277,7 @@ namespace ranges {
 
 		OMap make_edge_map(bool count_corners) {
 			OMap edge_map = OMap(width, height);
-			int occupied = 0;
+			// int occupied = 0;
 			for (int x = 0; x < width; ++x) {
 				for (int y = 0; y < height; ++y) {
 					if (!isOccupiedNT(x,y)) continue;
@@ -306,6 +313,7 @@ namespace ranges {
 		unsigned width;
 		unsigned height;
 		std::vector<std::vector<float> > grid;
+		// float *grid;
 
 		float get(int x, int y) { return grid[x][y]; }
 
@@ -325,17 +333,17 @@ namespace ranges {
 			width = map->width;
 			height = map->height;
 
[…]

### 2017-03-05 includes/RangeUtils.h

@@ -56,18 +56,18 @@ namespace utils {
 	public:
 		KeyMaker() {}
 		KeyMaker(int width, int height, int theta_discretization) {
-			y_shift = (int) ceil(log2(theta_discretization));
-			x_shift = (int) ceil(log2(height)) + y_shift;
-			int bitness = (int) ceil(log2(width)) + x_shift;
+			y_shift = (int) std::ceil(std::log2(theta_discretization));
+			x_shift = (int) std::ceil(std::log2(height)) + y_shift;
+			int bitness = (int) std::ceil(std::log2(width)) + x_shift;
 
-			if (bitness > log2(std::numeric_limits<key_T>::max())) {
+			if (bitness > std::log2(std::numeric_limits<key_T>::max())) {
 				std::cerr << "Key bitness too large for integer packing scheme. Check your KeyMaker template type." << std::endl;
 			}
 
 			// make bit masks for unpacking the various values
-			t_mask = pow(2, y_shift)-1;
-			y_mask = pow(2, x_shift)-1 - t_mask;
-			x_mask = pow(2, bitness)-1 - y_mask - t_mask;
+			t_mask = std::pow(2, y_shift)-1;
+			y_mask = std::pow(2, x_shift)-1 - t_mask;
+			x_mask = std::pow(2, bitness)-1 - y_mask - t_mask;
 		};
 		~KeyMaker() {};
 		key_T make_key(int x, int y, int t) {
@@ -236,8 +236,6 @@ namespace utils {
 		serialize(vals,&ss);
 		return ss.str();
 	}
-
-
 
 } // namespace utils
 

### 2017-03-05 includes/kernels.cu

@@ -1,0 +1,86 @@
+#include "includes/CudaRangeLib.h"
+// #include "includes/RangeUtils.h"
+
+
+#include <cuda.h>
+#include <string>
+#include <iostream>
+#include <sstream>
+#include <vector>
+
+#define DIST_THRESHOLD 0.0
+#define STEP_COEFF 0.999
+
+__device__ float distance(int x, int y, float *distMap, int width, int height) {
+	return distMap[x * height + y];
+}
+
+__global__ void cuda_ray_marching(float * ins, float * outs, float * distMap, int width, int height, float max_range, int num_casts) {
+	int ind = blockIdx.x*blockDim.x + threadIdx.x;
+	if (ind >= num_casts) return; 
+	float x0 = ins[ind*3];
+	float y0 = ins[ind*3+1];
+	float theta = ins[ind*3+2];
+
+	float ray_direction_x = cosf(theta);
+	float ray_direction_y = sinf(theta);
+
+	int px = 0;
+	int py = 0;
+
+	float t = 0.0;
+	float out = max_range;
+	// int iters = 0;
+	while (t < max_range) {
+		px = x0 + ray_direction_x * t;
+		py = y0 + ray_direction_y * t;
+
+		if (px >= width || px < 0 || py < 0 || py >= height) {
+			out = max_range;
+			break;
+		}
+
+		float d = distance(px,py, distMap, width, height);
+
+		if (d <= DIST_THRESHOLD) {
+			float xd = px - x0;
+			float yd = py - y0;
+			out =  sqrtf(xd*xd + yd*yd);
+			break;
+		}
+
+		t += fmaxf(d * STEP_COEFF, 1.0);
+		// iters ++;
+	}
+	outs[ind] = out;
+}
+
+RayMarchingCUDA::RayMarchingCUDA(std::vector<std::vector<float> > grid, int w, int h, float mr) 
+	: width(w), height(h), max_range(mr) {
+	cudaMalloc((void **)&d_ins, sizeof(float) * CHUNK_SIZE * 3);
+	cudaMalloc((void **)&d_outs, sizeof(float) * CHUNK_SIZE);
+	cudaMalloc((void **)&d_distMap, sizeof(float) * width * height);
+
+	// convert vector format to raw float array, y axis is quickly changing dimension
+	float raw_grid[width*height];
[…]

### 2017-03-05 main.cpp

@@ -57,7 +57,9 @@ DEFINE_string(lut_slice_theta, "1.57", "Which LUT slice to output");
 #define GRID_STEP 10
 #define GRID_RAYS 40
 #define GRID_SAMPLES 1
-#define RANDOM_SAMPLES 100
+#define RANDOM_SAMPLES 200000
+// #define RANDOM_SAMPLES (CHUNK_SIZE*2)
+// #define RANDOM_SAMPLES (CHUNK_SIZE*10)
 
 using namespace ranges;
 using namespace benchmark;
@@ -100,6 +102,10 @@ int main(int argc, char *argv[])
 	} else {
 		map = OMap(FLAGS_map_path);
 	}
+
+
+	// DistanceTransform dt = DistanceTransform(&map);
+	// dt.save("./test.png");
 
 	bool DO_LOG = (FLAGS_log_path != "");
 	std::stringstream tlog;
@@ -174,6 +180,7 @@ int main(int argc, char *argv[])
 		std::chrono::duration<double> construction_dur = 
 			std::chrono::duration_cast<std::chrono::duration<double>>(construction_end - construction_start);
 		Benchmark<RayMarching> mark = Benchmark<RayMarching>(rm);
+		std::cout << "...construction time: " << construction_dur.count() << std::endl;
 		std::cout << "...Running grid benchmark" << std::endl;
 		if (DO_LOG) {
 			tlog.str("");
@@ -303,6 +310,72 @@ int main(int argc, char *argv[])
 		if (DO_LOG) {
 			save_log(tlog, FLAGS_log_path+"/glt.csv");
 		}
+	}
+
+	if (utils::has("RayMarchingGPU", methods) || utils::has("rmgpu", methods)) {
+		#ifdef USE_CUDA
+		std::cout << "\n...Loading range method: RayMarchingGPU" << std::endl;
+		auto construction_start = std::chrono::high_resolution_clock::now();
+		RayMarchingGPU rmgpu = RayMarchingGPU(map, MAX_DISTANCE);
+		auto construction_end = std::chrono::high_resolution_clock::now();
+		std::chrono::duration<double> construction_dur = 
+			std::chrono::duration_cast<std::chrono::duration<double>>(construction_end - construction_start);
+		std::cout << "...construction time: " << construction_dur.count() << std::endl;
+
+		if (FLAGS_which_benchmark == "grid") {
[…]
